%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% AAMAS-2026 paper skeleton for the domain-level lifted ASL method.
%%% Template files are copied from the official AAMAS-2026 author kit.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[sigconf,anonymous]{aamas}

\usepackage{amsmath}
\usepackage{algorithm}
\usepackage{algpseudocode}

\makeatletter
\@ACM@balancefalse
\makeatother

\title[Goal-Conditioned Modular Policy Sketches]{Goal-Conditioned Modular
Policy Sketches for Lifted AgentSpeak Plan Libraries}

\author{Anonymous Authors}
\affiliation{
	\institution{Anonymous Institution}
	\city{Anonymous City}
	\country{Anonymous Country}}
\email{anonymous@example.org}

\setcopyright{ifaamas}
\acmConference[AAMAS '26]{25th International Conference
on Autonomous Agents and Multiagent Systems (AAMAS 2026)}{May 25 -- 29, 2026}
{Paphos, Cyprus}{C.~Amato, L.~Dennis, V.~Mascardi, J.~Thangarajah (eds.)}
\copyrightyear{2026}
\acmYear{2026}
\acmDOI{}
\acmPrice{}
\acmISBN{}
\acmSubmissionID{anonymous}

\begin{abstract}
We study the automatic synthesis of domain-level lifted AgentSpeak(L) plan
libraries from Planning Domain Definition Language domains and small sets of
achievement-goal training problems. The generated library is not a trace for a
single problem and is not tied to a query. Instead, its plan heads are lifted
predicate achievement goals and its bodies call declared primitive actions or
other predicate goals. The method combines goal-conditioned policy sketches,
modular predicate-goal rules, an answer-set selection problem, and a conservative
AgentSpeak(L) compiler. The claim is bounded to domains whose reusable strategy
is expressible as a goal-conditioned, feature-definable, terminating modular
policy sketch with safely bindable features. The current scope is positive
conjunctive achievement goals; temporal goals are handled by a separate
controller and are outside this paper core.
\end{abstract}

\ccsdesc[500]{Computing methodologies~Planning and scheduling}
\ccsdesc[300]{Computing methodologies~Multi-agent systems}

\keywords{generalized planning, policy sketches, AgentSpeak(L), plan libraries,
answer set programming, Planning Domain Definition Language}

\newcommand{\pddl}{Planning Domain Definition Language}
\newcommand{\asl}{AgentSpeak(L)}
\newcommand{\goalpred}[1]{\ensuremath{\mathsf{goal}\_#1}}
\newcommand{\select}{\ensuremath{\mathsf{select}}}
\newcommand{\compile}{\ensuremath{\mathsf{compile}}}
\newcommand{\methodname}{Goal-Conditioned Modular Policy Sketches}
% BEGIN GENERATED RESULT MACROS
% Auto-generated by scripts/compare_domain_level_experiments.py.
% Re-run the comparison script after changing experiment reports.
\newcommand{\ResultPaperExternalSketchFirstTwoZeroType}{library}
\newcommand{\ResultPaperExternalSketchFirstTwoZeroSolved}{20/20}
\newcommand{\ResultPaperExternalSketchFirstTwoZeroCoveragePercent}{100.0\%}
\newcommand{\ResultPaperExternalSketchFirstTwoZeroRuntimePlanner}{none}
\newcommand{\ResultPaperExternalSketchFirstTwoZeroMechanisms}{enabled: external-sketch-evidence, layer-c-ordering, paper-profile-gate}
\newcommand{\ResultPaperExternalSketchFirstTwoZeroPlanCount}{41}
\newcommand{\ResultPaperExternalSketchFirstTwoZeroNotes}{selected external sketch evidence}
\newcommand{\ResultPaperExternalSketchSatisfiableLargeType}{library}
\newcommand{\ResultPaperExternalSketchSatisfiableLargeSolved}{10/10}
\newcommand{\ResultPaperExternalSketchSatisfiableLargeCoveragePercent}{100.0\%}
\newcommand{\ResultPaperExternalSketchSatisfiableLargeRuntimePlanner}{none}
\newcommand{\ResultPaperExternalSketchSatisfiableLargeMechanisms}{enabled: external-sketch-evidence, layer-c-ordering, paper-profile-gate}
\newcommand{\ResultPaperExternalSketchSatisfiableLargePlanCount}{41}
\newcommand{\ResultPaperExternalSketchSatisfiableLargeNotes}{selected external sketch evidence}
\newcommand{\ResultPaperExternalSketchSatisfiableMixedLargeType}{library}
\newcommand{\ResultPaperExternalSketchSatisfiableMixedLargeSolved}{10/10}
\newcommand{\ResultPaperExternalSketchSatisfiableMixedLargeCoveragePercent}{100.0\%}
\newcommand{\ResultPaperExternalSketchSatisfiableMixedLargeRuntimePlanner}{none}
\newcommand{\ResultPaperExternalSketchSatisfiableMixedLargeMechanisms}{enabled: external-sketch-evidence, layer-c-ordering, paper-profile-gate}
\newcommand{\ResultPaperExternalSketchSatisfiableMixedLargePlanCount}{41}
\newcommand{\ResultPaperExternalSketchSatisfiableMixedLargeNotes}{selected external sketch evidence}
\newcommand{\ResultLabworkflowCounterexamplHashncinkmnmType}{library}
\newcommand{\ResultLabworkflowCounterexamplHashncinkmnmSolved}{5/5}
\newcommand{\ResultLabworkflowCounterexamplHashncinkmnmCoveragePercent}{100.0\%}
\newcommand{\ResultLabworkflowCounterexamplHashncinkmnmRuntimePlanner}{none}
\newcommand{\ResultLabworkflowCounterexamplHashncinkmnmMechanisms}{enabled: counterexample-refinement, layer-c-ordering}
\newcommand{\ResultLabworkflowCounterexamplHashncinkmnmPlanCount}{11}
\newcommand{\ResultLabworkflowCounterexamplHashncinkmnmNotes}{}
\newcommand{\ResultNoExternalSketchFirstTwoZeroType}{library}
\newcommand{\ResultNoExternalSketchFirstTwoZeroSolved}{20/20}
\newcommand{\ResultNoExternalSketchFirstTwoZeroCoveragePercent}{100.0\%}
\newcommand{\ResultNoExternalSketchFirstTwoZeroRuntimePlanner}{none}
\newcommand{\ResultNoExternalSketchFirstTwoZeroMechanisms}{enabled: layer-c-ordering}
\newcommand{\ResultNoExternalSketchFirstTwoZeroPlanCount}{40}
\newcommand{\ResultNoExternalSketchFirstTwoZeroNotes}{schema-only bootstrap}
\newcommand{\ResultNoLayerCWithRefinementLabworkflowStressType}{library}
\newcommand{\ResultNoLayerCWithRefinementLabworkflowStressSolved}{0/5}
\newcommand{\ResultNoLayerCWithRefinementLabworkflowStressCoveragePercent}{0.0\%}
\newcommand{\ResultNoLayerCWithRefinementLabworkflowStressRuntimePlanner}{none}
\newcommand{\ResultNoLayerCWithRefinementLabworkflowStressMechanisms}{enabled: counterexample-refinement}
\newcommand{\ResultNoLayerCWithRefinementLabworkflowStressPlanCount}{10}
\newcommand{\ResultNoLayerCWithRefinementLabworkflowStressNotes}{no applicable plan for !reagent-logged(r1): 5}
\newcommand{\ResultNoCounterexampleRefinemeHashamkpcpajType}{library}
\newcommand{\ResultNoCounterexampleRefinemeHashamkpcpajSolved}{0/5}
\newcommand{\ResultNoCounterexampleRefinemeHashamkpcpajCoveragePercent}{0.0\%}
\newcommand{\ResultNoCounterexampleRefinemeHashamkpcpajRuntimePlanner}{none}
\newcommand{\ResultNoCounterexampleRefinemeHashamkpcpajMechanisms}{enabled: layer-c-ordering}
\newcommand{\ResultNoCounterexampleRefinemeHashamkpcpajPlanCount}{10}
\newcommand{\ResultNoCounterexampleRefinemeHashamkpcpajNotes}{no applicable plan for !reagent-logged(r1): 5}
\newcommand{\ResultNoOfflineTraceEvidenceTransportType}{library}
\newcommand{\ResultNoOfflineTraceEvidenceTransportSolved}{10/10}
\newcommand{\ResultNoOfflineTraceEvidenceTransportCoveragePercent}{100.0\%}
\newcommand{\ResultNoOfflineTraceEvidenceTransportRuntimePlanner}{none}
\newcommand{\ResultNoOfflineTraceEvidenceTransportMechanisms}{enabled: layer-c-ordering}
\newcommand{\ResultNoOfflineTraceEvidenceTransportPlanCount}{32}
\newcommand{\ResultNoOfflineTraceEvidenceTransportNotes}{schema-only bootstrap}
\newcommand{\ResultTransportInitialStateFragmentType}{library}
\newcommand{\ResultTransportInitialStateFragmentSolved}{10/10}
\newcommand{\ResultTransportInitialStateFragmentCoveragePercent}{100.0\%}
\newcommand{\ResultTransportInitialStateFragmentRuntimePlanner}{none}
\newcommand{\ResultTransportInitialStateFragmentMechanisms}{enabled: offline-synthesis-planner-traces, layer-c-ordering}
\newcommand{\ResultTransportInitialStateFragmentPlanCount}{32}
\newcommand{\ResultTransportInitialStateFragmentNotes}{schema-only bootstrap}
\newcommand{\ResultMarsroverTraceEvidenceFragmentType}{library}
\newcommand{\ResultMarsroverTraceEvidenceFragmentSolved}{10/10}
\newcommand{\ResultMarsroverTraceEvidenceFragmentCoveragePercent}{100.0\%}
\newcommand{\ResultMarsroverTraceEvidenceFragmentRuntimePlanner}{none}
\newcommand{\ResultMarsroverTraceEvidenceFragmentMechanisms}{enabled: offline-synthesis-planner-traces, layer-c-ordering}
\newcommand{\ResultMarsroverTraceEvidenceFragmentPlanCount}{123}
\newcommand{\ResultMarsroverTraceEvidenceFragmentNotes}{schema-only bootstrap}
\newcommand{\ResultBlocksworldPaperExternalHashjkajkcmaType}{baseline}
\newcommand{\ResultBlocksworldPaperExternalHashjkajkcmaSolved}{20/20}
\newcommand{\ResultBlocksworldPaperExternalHashjkajkcmaCoveragePercent}{100.0\%}
\newcommand{\ResultBlocksworldPaperExternalHashjkajkcmaRuntimePlanner}{offline_baseline_only}
\newcommand{\ResultBlocksworldPaperExternalHashjkajkcmaMechanisms}{baseline}
\newcommand{\ResultBlocksworldPaperExternalHashjkajkcmaPlanCount}{--}
\newcommand{\ResultBlocksworldPaperExternalHashjkajkcmaNotes}{per-problem-trace-baseline; executed-and-strips-validated; solves each evaluation problem independently; not a domain-level library}
\newcommand{\ResultBlocksworldPaperExternalHashkkicclnjType}{baseline}
\newcommand{\ResultBlocksworldPaperExternalHashkkicclnjSolved}{0/20}
\newcommand{\ResultBlocksworldPaperExternalHashkkicclnjCoveragePercent}{0.0\%}
\newcommand{\ResultBlocksworldPaperExternalHashkkicclnjRuntimePlanner}{not_runtime_executed}
\newcommand{\ResultBlocksworldPaperExternalHashkkicclnjMechanisms}{baseline}
\newcommand{\ResultBlocksworldPaperExternalHashkkicclnjPlanCount}{--}
\newcommand{\ResultBlocksworldPaperExternalHashkkicclnjNotes}{artifact-audit; not-directly-executable-without-asl-binding; domain-level; raw qualitative sketch audit; ready-for-executable-asl=True}
\newcommand{\ResultBlocksworldPaperExternalHashldfkjnbdType}{baseline}
\newcommand{\ResultBlocksworldPaperExternalHashldfkjnbdSolved}{0/5}
\newcommand{\ResultBlocksworldPaperExternalHashldfkjnbdCoveragePercent}{0.0\%}
\newcommand{\ResultBlocksworldPaperExternalHashldfkjnbdRuntimePlanner}{moose_policy_execution_or_training_probe}
\newcommand{\ResultBlocksworldPaperExternalHashldfkjnbdMechanisms}{baseline}
\newcommand{\ResultBlocksworldPaperExternalHashldfkjnbdPlanCount}{--}
\newcommand{\ResultBlocksworldPaperExternalHashldfkjnbdNotes}{generalized-planning-reproduction; completed-status-csv; domain-level; imported from completed MOOSE reproduction status}
\newcommand{\ResultBlocksworldPaperExternalHashfcjedmihType}{baseline}
\newcommand{\ResultBlocksworldPaperExternalHashfcjedmihSolved}{0/10}
\newcommand{\ResultBlocksworldPaperExternalHashfcjedmihCoveragePercent}{0.0\%}
\newcommand{\ResultBlocksworldPaperExternalHashfcjedmihRuntimePlanner}{offline_baseline_only}
\newcommand{\ResultBlocksworldPaperExternalHashfcjedmihMechanisms}{baseline}
\newcommand{\ResultBlocksworldPaperExternalHashfcjedmihPlanCount}{--}
\newcommand{\ResultBlocksworldPaperExternalHashfcjedmihNotes}{per-problem-trace-baseline; executed-and-strips-validated; solves each evaluation problem independently; not a domain-level library}
\newcommand{\ResultBlocksworldPaperExternalHashakcbodedType}{baseline}
\newcommand{\ResultBlocksworldPaperExternalHashakcbodedSolved}{0/10}
\newcommand{\ResultBlocksworldPaperExternalHashakcbodedCoveragePercent}{0.0\%}
\newcommand{\ResultBlocksworldPaperExternalHashakcbodedRuntimePlanner}{not_runtime_executed}
\newcommand{\ResultBlocksworldPaperExternalHashakcbodedMechanisms}{baseline}
\newcommand{\ResultBlocksworldPaperExternalHashakcbodedPlanCount}{--}
\newcommand{\ResultBlocksworldPaperExternalHashakcbodedNotes}{artifact-audit; not-directly-executable-without-asl-binding; domain-level; raw qualitative sketch audit; ready-for-executable-asl=True}
\newcommand{\ResultBlocksworldPaperExternalHashbaeacebfType}{baseline}
\newcommand{\ResultBlocksworldPaperExternalHashbaeacebfSolved}{0/5}
\newcommand{\ResultBlocksworldPaperExternalHashbaeacebfCoveragePercent}{0.0\%}
\newcommand{\ResultBlocksworldPaperExternalHashbaeacebfRuntimePlanner}{moose_policy_execution_or_training_probe}
\newcommand{\ResultBlocksworldPaperExternalHashbaeacebfMechanisms}{baseline}
\newcommand{\ResultBlocksworldPaperExternalHashbaeacebfPlanCount}{--}
\newcommand{\ResultBlocksworldPaperExternalHashbaeacebfNotes}{generalized-planning-reproduction; completed-status-csv; domain-level; imported from completed MOOSE reproduction status}
\newcommand{\ResultBlocksworldPaperExternalHashkgepfcfeType}{baseline}
\newcommand{\ResultBlocksworldPaperExternalHashkgepfcfeSolved}{7/10}
\newcommand{\ResultBlocksworldPaperExternalHashkgepfcfeCoveragePercent}{70.0\%}
\newcommand{\ResultBlocksworldPaperExternalHashkgepfcfeRuntimePlanner}{offline_baseline_only}
\newcommand{\ResultBlocksworldPaperExternalHashkgepfcfeMechanisms}{baseline}
\newcommand{\ResultBlocksworldPaperExternalHashkgepfcfePlanCount}{--}
\newcommand{\ResultBlocksworldPaperExternalHashkgepfcfeNotes}{per-problem-trace-baseline; executed-and-strips-validated; solves each evaluation problem independently; not a domain-level library}
\newcommand{\ResultBlocksworldPaperExternalHashaeekomkeType}{baseline}
\newcommand{\ResultBlocksworldPaperExternalHashaeekomkeSolved}{0/10}
\newcommand{\ResultBlocksworldPaperExternalHashaeekomkeCoveragePercent}{0.0\%}
\newcommand{\ResultBlocksworldPaperExternalHashaeekomkeRuntimePlanner}{not_runtime_executed}
\newcommand{\ResultBlocksworldPaperExternalHashaeekomkeMechanisms}{baseline}
\newcommand{\ResultBlocksworldPaperExternalHashaeekomkePlanCount}{--}
\newcommand{\ResultBlocksworldPaperExternalHashaeekomkeNotes}{artifact-audit; not-directly-executable-without-asl-binding; domain-level; raw qualitative sketch audit; ready-for-executable-asl=True}
\newcommand{\ResultBlocksworldPaperExternalHashbbgdmpblType}{baseline}
\newcommand{\ResultBlocksworldPaperExternalHashbbgdmpblSolved}{0/5}
\newcommand{\ResultBlocksworldPaperExternalHashbbgdmpblCoveragePercent}{0.0\%}
\newcommand{\ResultBlocksworldPaperExternalHashbbgdmpblRuntimePlanner}{moose_policy_execution_or_training_probe}
\newcommand{\ResultBlocksworldPaperExternalHashbbgdmpblMechanisms}{baseline}
\newcommand{\ResultBlocksworldPaperExternalHashbbgdmpblPlanCount}{--}
\newcommand{\ResultBlocksworldPaperExternalHashbbgdmpblNotes}{generalized-planning-reproduction; completed-status-csv; domain-level; imported from completed MOOSE reproduction status}
\newcommand{\ResultLabworkflowCounterexamplHashflonfhglType}{baseline}
\newcommand{\ResultLabworkflowCounterexamplHashflonfhglSolved}{5/5}
\newcommand{\ResultLabworkflowCounterexamplHashflonfhglCoveragePercent}{100.0\%}
\newcommand{\ResultLabworkflowCounterexamplHashflonfhglRuntimePlanner}{offline_baseline_only}
\newcommand{\ResultLabworkflowCounterexamplHashflonfhglMechanisms}{baseline}
\newcommand{\ResultLabworkflowCounterexamplHashflonfhglPlanCount}{--}
\newcommand{\ResultLabworkflowCounterexamplHashflonfhglNotes}{per-problem-trace-baseline; executed-and-strips-validated; solves each evaluation problem independently; not a domain-level library}
% END GENERATED RESULT MACROS

\begin{document}

\pagestyle{fancy}
\fancyhead{}

\maketitle

\section{Introduction}

Classical planners can generate a plan for a concrete problem, but a
Belief-Desire-Intention agent requires a reusable plan library. AgentSpeak(L)
provides a logical language for such agents \cite{Rao1996AgentSpeak}. This
paper targets the gap between generalized planning and executable agent
programs: given a \pddl{} domain and a small set of training problems,
synthesize a lifted \asl{} library that can solve new achievement-goal problems
from the same domain without asking a planner for a full trace at run time.
The resulting artifact is ordinary agent code: a finite set of lifted plan
schemas whose triggers, contexts, and bodies can be inspected independently of
any particular test instance.

The central difficulty is goal dependence. A domain-level library cannot simply
learn one trace per problem, nor can it assume that conjunctive goals are
independent. MOOSE formalizes useful goal-independence assumptions
\cite{Chen2026Moose}, and sketches formalize goal serialization and
bounded-width subgoal structure \cite{Drexler2024SubgoalSketches,
Bonet2024GeneralPoliciesWidth}. We use these established boundaries rather
than introducing a new domain class. In Blocksworld, for example, a tower goal
must be built in a causally valid order; achieving a top block before its
support can destroy the conditions needed later. The library must therefore
contain both atomic predicate-goal modules and a composer that chooses which
goal to pursue next.

We build on the observation that many planning domains share compact subgoal
structure expressible by policy sketches and generalized policies
\cite{Drexler2022LearningSketches,Drexler2024SubgoalSketches,
Bonet2024GeneralPoliciesWidth}. Our contribution is to turn that structure into
an executable \asl{} artifact rather than a planner wrapper. External sketches
provide qualitative evidence, \pddl{} schemas and traces provide lifted support
evidence, and an answer-set selector chooses only rules that the compiler can
render safely. The output is not a per-problem plan and does not contain
synthetic achievement names. Atomic modules implement declared \pddl{}
predicates such as \texttt{clear} or \texttt{on}. A separate composer uses
read-only goal descriptors such as \texttt{goal\_on} and derived
\texttt{ready\_on} contexts to order conjunctive goals. The descriptors say what
the current problem asks for; the achievement-goal heads remain ordinary
\pddl{} predicates.

The paper makes four contributions.
\begin{enumerate}
	\item We define a bounded hypothesis class for goal-conditioned modular
	policy sketches that compile to lifted \asl{} plan schemas, together with a
	literature-backed domain-support taxonomy and explicit exclusions for
	unsupported \pddl{} fragments.
	\item We give an answer-set selection procedure that chooses Layer B
	atomic modules and Layer C composer rules from \pddl{} schemas, verified
	external sketches, bounded transition evidence, and lifted counterexamples.
	\item We provide a conservative feature-binding and compiler contract:
	every external feature must map to declared predicates, primitive actions,
	predicate subgoals, read-only goal descriptors, or a structured rejection.
	\item We evaluate the resulting libraries on achievement-goal domains with
	no run-time full-trace planner, including strict Blocksworld paper-profile
	rows and diagnostic ablations for refinement and goal composition.
\end{enumerate}

The contribution is deliberately bounded. We do not claim universal
generalized-planning completeness for arbitrary \pddl{} domains. The main
positive family in this paper is goal-dependent, feature-definable
Blocksworld-style achievement goals. Ferry and Gripper are natural control
candidates; Spanner and Childsnack are the next ordered-resource stress
candidates. Transport and Marsrover remain boundary rows until richer route and
ranking features are accepted. Unsupported cases are rejected with diagnostics
rather than compiled into unsafe plans.

\input{sections/related_work}

\input{sections/method}

\input{sections/evaluation}

\section{Current Scope and Exclusions}
\label{sec:scope}

The current method supports positive conjunctive achievement goals only for the
bounded domain family defined in Section~\ref{sec:method}: the reusable strategy
must be expressible as a goal-conditioned, feature-definable, terminating
modular sketch with safe lifted feature bindings. Negative goals, disjunctive
goals, derived predicates, conditional effects, and numeric fluents require
additional semantics before they can be part of the paper claim. Unsupported
cases are rejected before compilation so that the generated library never
contains undeclared symbols, grounded training objects, or synthetic
achievement names.

The feature language is also bounded. The compiler accepts only external
features with recoverable lifted semantics. Object-free goal-distance features
are supported when they refer to a declared unary state predicate, the matching
read-only goal predicate, and a declared binary role. Concrete object-specific
features remain outside the claim because they would make the output depend on
training-instance object names.

Temporal extended goals are handled by a future deterministic finite automaton
controller that calls the achievement-goal library; they are not part of the
current algorithm. This separation is intentional: the deterministic finite
automaton layer is query-specific, while the achievement-goal library remains
domain-level.

\bibliographystyle{ACM-Reference-Format}
\bibliography{references}

\end{document}
