\section{Problem Formulation and Foundations}
\label{sec:background}

\subsection{Planning Domains and Generalized Evidence}

We use typed STRIPS PDDL plus a bounded-integer resource fragment
\cite{FikesNilsson1971STRIPS,McDermott1998PDDL,FoxLong2003PDDL21}. A domain is
$D=\langle\mathcal{P},\mathcal{F},\mathcal{A},\mathcal{T}\rangle$, where
$\mathcal{P}$ and $\mathcal{F}$ contain predicate schemas and integer-valued
resource functions, $\mathcal{A}$ contains action schemas, and $\mathcal{T}$ is
a type hierarchy. A state $s=(s_{\mathcal P},s_{\mathcal F})$ combines Boolean
and integer valuations. For an applicable grounded action $a$, the Boolean
successor is
\[
\gamma_{\mathcal P}(s,a)=
(s_{\mathcal P}\setminus\mathit{del}(a))\cup\mathit{add}(a).
\]
The numeric component admits finite integer values, comparisons with constants,
and constant integer increases or decreases; arbitrary arithmetic and
continuous change are excluded. A problem instance is
$I=\langle D,O_I,s_I^0,G_I\rangle$, where $O_I$ is its finite typed object set,
$s_I^0$ its initial state, and $G_I$ its PDDL achievement condition. A
generalized-planning task supplies finite training and test instance sets that
share $D$. The formal interface is agnostic to whether those finite sets are
curated or produced by a parameterized problem generator; their provenance and
separation remain part of the experimental protocol. MOOSE applies classical
goal-regression ideas
\cite{Fikes1972GeneralizedRobotPlans,Alcazar2013Regression} to singleton
goal conditions and
lifts the resulting plans into first-order condition-to-macro rules
\cite{Chen2026Moose}. We treat these rules as evidence: they demonstrate
ways to achieve observed singleton goals, but need not constitute a closed or
compact atomic module core.

\subsection{BDI Plan Libraries and Their AgentSpeak(L) Realization}

A procedural BDI plan rule has a triggering event, an applicability context,
and a body \cite{MeneguzziDeSilva2015PlanningBDI}. In the \asl{} realization
evaluated here \cite{Rao1996AgentSpeak}, an achievement plan is
\[
+!p(\bar X):C\leftarrow b_1;\ldots;b_m,
\]
where $C$ is a conjunction of beliefs and comparisons and each $b_i$ is a PDDL
action or an internal achievement call $!q(\bar Y)$. A library is lifted when
plans contain variables rather than training objects, and internally closed
when every internal call resolves to a type-compatible selected module.
Internal-call closure is a signature-resolution property; it does not assert
that some plan for the call is applicable in every reachable state.

Our certificates apply to this trigger--context--body model and to the PDDL
semantics of primitive actions. Formally, a candidate branch is
$b=\langle g_b,C_b,\beta_b,\Sigma_b,\pi_b\rangle$, where $g_b$ is one lifted
achievement literal, $C_b$ is a conjunctive applicability condition, $\beta_b$
is a finite body of primitive actions and internal calls, $\Sigma_b$ is a
conservative conditional module-completion summary, and $\pi_b$ records branch
provenance. We map selected
branches to \asl{} and evaluate them with Jason \cite{Bordini2007Jason}; claims
about another BDI language would require a semantics-preserving translation and
an independent execution study. Query controllers follow declarative goal
patterns by rechecking their intended state after subsidiary plans return
\cite{Hubner2006DeclarativeGoals}.

\subsection{Finite-Trace Temporal Goals}

Linear temporal logic over finite traces (\ltlf{}) is interpreted over a
nonempty finite state sequence \cite{DeGiacomoVardi2013}. For proposition
alphabet $AP_q$, the declared input language $\Phi_{\mathrm{syn}}$ admits
literals, conjunction, $F$, $X$, and strong $U$ over $AP_q$; the full grammar
and finite-trace semantics appear in the Technical Supplement, Sec.~1.
Negation is restricted to atoms; the serialized syntax writes it as
\texttt{!}. Both $X$ and $U$ have their strong finite-trace meanings. The
evaluation family $\Phi_{\mathrm{bench}}\subset\Phi_{\mathrm{syn}}$ contains
instances of the five predeclared profiles in Sec.~\ref{sec:evaluation}, and
the certificate-supported subset $\Phi_{\mathrm{cert}}$ is defined below.
Membership in either set is not an action-strategy guarantee.

Automata-theoretic planning uses a finite automaton to track temporal progress
\cite{DeGiacomoRubin2018AutomataFOND}. We construct the deterministic finite
automaton (DFA)
$\mathcal D_q=\langle Q_q^{\mathrm{dfa}},2^{AP_q},\delta_q,q_q^0,F_q\rangle$
with LTLf2DFA and MONA
\cite{Fuggitti2020LTLf2DFA,Henriksen1995MONA}. Here
$Q_q^{\mathrm{dfa}}$ is the finite state set, $AP_q$ the proposition alphabet,
$\delta_q$ the deterministic transition function, $q_q^0$ the initial state,
and $F_q$ the accepting set. The valuation map
$\mathit{val}_q:\operatorname{States}(D)\rightarrow 2^{AP_q}$ interprets a PDDL
state over that alphabet. The compiler-supported subset requires certified
progress transitions. For automaton state $z$, let
$d_q(z)$ be the shortest directed-path length to $F_q$, or $\infty$ when no
accepting state is reachable. A progress transition $(q_s,\chi,q_t)$ satisfies
$d_q(q_t)<d_q(q_s)<\infty$ and has a guard containing
conjunction and literal negation. A supported guard $\chi$ has normal form
\[
\chi=
\bigwedge_{G\in P_\chi^+}G\ \land\!
\bigwedge_{N\in P_\chi^-}\neg N,
\qquad
\mathcal O_\chi=\langle P_\chi^+,P_\chi^-\rangle,
\]
where the signed transition obligation $\mathcal O_\chi$ separates positive
achievements $P_\chi^+$ from atoms $P_\chi^-$ that must remain absent. MONA may
represent Boolean alternatives by several valuation cubes. GP2PL
treats each cube as a separate conjunctive guard; several cubes are not one
disjunctive guard, and the monitor continues to evaluate every original cube.
An explicit disjunction inside one guard and strategy choice without a
certificate remain unsupported. The certificate-dependent subset
$\Phi_{\mathrm{cert}}(D,\mathcal M_D)\subseteq\Phi_{\mathrm{syn}}$ contains the
validated bound formulas whose required DFA progress obligations admit the
certificates below. The soundness claims apply to this subset; the empirical
claims apply to accepted members of $\Phi_{\mathrm{bench}}$. A query-local
monitor evaluates the DFA initially and after every successful primitive PDDL
action. Arbitrary arithmetic and progress without a certified action strategy
are outside this fragment.

\subsection{Certified Compilation Problem}

\textbf{Compilation problem.} Given $D$, training instances
$\mathcal{I}_{\mathrm{train}}$, and normalized singleton-goal evidence $E$,
atomic compilation produces the certified atomic module core $\mathcal{M}_D$:
the query-independent branches whose triggers are singleton achievements.

An unbound temporal specification is $\tau_q$, its externally supplied
type-consistent object binding is $\theta_q$, and
$\widehat\tau_q=(\tau_q,\theta_q)$ is the validated bound query. Temporal
compilation produces the query-local plan set $\mathcal{Q}_q$, comprising its
top-level goal, DFA dispatch, and transition-repair plans. For appended queries
$q_1,\ldots,q_k$, the one maintained domain library is
\[
\mathcal{L}_D^{[0]}=\mathcal{M}_D,
\qquad
\mathcal{L}_D^{[k]}=
\mathcal{M}_D\cup\bigcup_{i=1}^{k}\mathcal{Q}_{q_i}.
\]
Thus $\mathcal M_D$ is the stable logical core of $\mathcal L_D^{[k]}$, not a
second persisted library. Throughout, calligraphic symbols denote finite sets or
structured artifacts; $b$ denotes one candidate branch, $e$ one normalized
evidence rule, and $\chi$ one DFA transition guard. The Technical Supplement,
Sec.~1 defines the generated candidate set, certificates, preservation
portfolios, and transition-repair programs used below.

Positive predicate goals and supported integer equalities require an executable
achieving module. Negative literals denote absence and preservation, never
synthetic negative achievement calls. Conjunctive guards require a
preservation-safe order; mixed Boolean--numeric guards require complete net
effects for a common final state. Compilation rejects unsupported arithmetic,
disjunction, cyclic threats without a preserving strategy, and any obligation
whose effects are incomplete. In the \asl{} realization, static sort-membership
beliefs may constrain PDDL type compatibility, but no domain-specific typing
predicate becomes an achievement goal or primitive action. Formal definitions
and supported-fragment assumptions appear in the Technical Supplement, Sec.~1.
