\section{Experimental Evaluation}
\label{sec:evaluation}

\paragraph{Questions.}
The evaluation isolates the two compilation contributions before reporting
end-to-end execution. RQ1 asks whether schema-certified lifting improves
module closure and held-out execution over validated direct adaptation of the
same evidence. RQ2 measures the separate contributions of schema closure,
decomposed candidate generation, and joint Clingo selection. RQ3 asks whether
completion-effect certificates improve temporal execution over a controller
that uses the same real DFA and atomic library without preservation reasoning.
RQ4 measures the structural effect of balanced transition controllers. RQ5
reports end-to-end validity, cost, and amortization relative to external
generalized- and instance-planning references.

\paragraph{Comparison principle.}
Our method is a representation compiler rather than a new evidence learner or
instance planner. The primary comparisons therefore hold the normalized
evidence, PDDL schemas, DFA, atomic library, and Jason runtime fixed and replace
only the compiler mechanism under study. This follows the controlled
same-representation comparisons used for generalized-policy scoring
\cite{Yang2022PG3} and the fixed-downstream-planner comparison of temporal
compilations in Plan4Past \cite{Bonassi2023Plan4Past}. Raw MOOSE, LAMA, and
ENHSP are reported separately as external task-level references; they cannot
establish which part of our compiler is responsible for an improvement.

\paragraph{Atomic compiler baselines and ablations.}
The strongest atomic baseline, \emph{Evidence Adapter}, symbolically validates
every provider macro against the PDDL action schemas
and renders the surviving lifted rules as AgentSpeak, but performs no
producible-fluent closure, internal-module synthesis, or branch optimization.
It answers why validated generalized-planning evidence cannot simply be
adapted to AgentSpeak. \emph{Action Closure} adds action-only schema closure.
\emph{Maximal Certified} additionally generates recursive preparation and
resource-discharge candidates, then retains a largest jointly compatible
program under the hard certificate constraints. This is not the unchecked
union of individually valid branches: recursive rankings and dependency
closure can make branches mutually incompatible. \emph{Full Compiler} applies
the compact Clingo objective to the same candidate universe and hard
constraints. Consequently all four methods use the same evidence artifact,
and Maximal Certified versus Full Compiler changes only the optimization
objective. These four cumulative rows are the complete registered atomic
matrix. We test individual progress, resource, preparation, and binding
certificate boundaries in the fail-closed challenge matrix rather than
disabling a certificate and emitting an unchecked branch.

\paragraph{Temporal compiler baselines and ablations.}
\emph{Unprotected DFA} uses the same real
MONA-derived DFA and primitive-step runtime monitor as the full method, but
serializes each conjunction in a deterministic canonical order without threat
ordering or preservation-safe branch portfolios. \emph{Certified Flat} adds
the complete effect-based ordering and portfolio certificates while retaining
a flat sibling controller. \emph{Certified Balanced} replaces only that
control structure with the balanced binary repair tree. \emph{Completion
Monitor} observes the DFA only when an atomic module returns, rather than after
every primitive action. It retains completion-effect certification and balanced
control but omits primitive-prefix source-invariant filtering because those
intermediate states are outside its observation semantics. It runs on the same
complete paired benchmark, while observation-boundary effects are attributed
only on formulas with intermediate-state obligations. These four rows are the
complete registered temporal matrix. Signed-negative and bounded-numeric cases
remain in the full benchmark with explicit support and failure statuses; we do
not report unimplemented capability-switch rows. The removed historical
sequence-only wrapper is not a production path; if reported, it is an
evaluation-only weak reference.

\begin{table*}[t]
\centering
\small
\caption{Registered primary baselines and ablations. Empty cells are populated
only from completed, hash-locked runs; no diagnostic value is inserted by
hand. Valid denotes the relevant VAL and DFA oracle, coverage is paired
held-out success, size is branches or controller plans, and time is PAR-2 or
the declared compilation cost.}
\label{tab:registered-ablations}
\begin{tabular}{lllrrrr}
\toprule
Stage & Method & Difference & Valid & Coverage & Size & Time \\
\midrule
Atomic & Evidence Adapter & Evidence macros only & -- & -- & -- & -- \\
Atomic & Action Closure & Add schema producers & -- & -- & -- & -- \\
Atomic & Maximal Certified & Maximize compatible branches & -- & -- & -- & -- \\
Atomic & Full Compiler & Minimize certified program & -- & -- & -- & -- \\
Temporal & Unprotected DFA & Canonical unprotected order & -- & -- & -- & -- \\
Temporal & Certified Flat & Threat-safe flat control & -- & -- & -- & -- \\
Temporal & Certified Balanced & Add balanced repair tree & -- & -- & -- & -- \\
Temporal & Completion Monitor & Observe module completions & -- & -- & -- & -- \\
\bottomrule
\end{tabular}
\end{table*}

\paragraph{External references.}
The exact MOOSE executor is run on the same learned model to distinguish
evidence quality from downstream compilation and AgentSpeak execution. LAMA is
the classical per-instance reference, and ENHSP MRP+HJ is the numeric
per-instance reference, matching the roles used in the MOOSE evaluation
\cite{Chen2026Moose}. The direct temporal reference uses the official
FOND4LTLf v0.0.4 compilation \cite{DeGiacomoFuggitti2021FOND4LTLf}, then invokes
the same fixed LAMA backend. It is evaluated only on its accepted grounded
Boolean subset; numeric PDDL and identifiers incompatible with the tool's atom
encoding are reported as unsupported rather than failed. Plan4Past is not
treated as directly interchangeable: it accepts pure-past formulas, and any
future-to-past translation would first require exact DFA-language equivalence.
Published numbers from different splits or hardware are related-work context,
not entries in our paired result tables.

\paragraph{Corpus and registered protocol.}
The corpus contains all eight MOOSE companion classical domains with nonempty
official train/test splits \cite{Chen2026Moose}, four bounded-integer variants,
and four serialized-width families. Blocksworld Clear and On use the published
no-constants general-policy splits \cite{Frances2021LearningGeneralPolicies};
Tower uses the first quarter of a pinned public PDDL corpus and Depots the first
half of the D2L corpus \cite{Bonet2018Features}. These groups describe
evaluation properties and never route compilation.

The registered atomic protocol uses the complete train split, singleton goals,
three sampled goal orders, and fixed seeds 0--4. The five repetitions estimate
sensitivity to MOOSE's randomized goal-order sampling. Each uses one internal
MOOSE worker so training problems consume the seeded permutation stream
sequentially rather than through schedule-dependent concurrent calls. Policies
are compiled and evaluated separately, never pooled or selected post hoc. We
report every repetition together with the mean and sample standard deviation.
Each synthesis run has a 12-hour timeout and 16-GiB memory guard;
Jason and VAL use 30-minute limits and six per-test workers. Independent seeds
may be launched concurrently to reduce elapsed wall time, but contended timing
is excluded from efficiency claims. MOOSE reports a 32-GB synthesis budget and
five repetitions \cite{Chen2026Moose}; our lower memory guard and deterministic
single-worker setting are declared reproduction choices.

The separate pinned temporal experiment \texttt{\TEGExecutionRunID} uses clean commit
\texttt{\TEGExecutionCommit}, 12 workers, 30-minute Jason and VAL limits, and a
64-MiB Java stack. It records SHA-256 hashes of every atomic JSON/ASL input. The
result generator rejects a changed benchmark hash, prediction hash, dirty
tracked revision, missing case, or mismatched library hash.

Every atomic comparison is paired by the SHA-256 hash of its exact normalized
evidence program. Temporal variants are paired by atomic-library, lifted-
\ltlf JSON, binding, and DFA hashes. The registered harness invokes production
compiler and execution paths rather than reimplementing them. It rejects an
atomic comparison unless every method shares the readable-policy and normalized-
evidence hashes, and rejects a temporal comparison unless every method shares
the benchmark, atomic-library, sample-set, and DFA hashes. Missing runs and
pairing violations remain explicit infrastructure failures. Evidence synthesis,
domain compilation, query append, execution, and validation are timed separately.
The DFA hash covers the formula, atom binding, initial and accepting states, and
guarded transition graph; it excludes conversion timing, artifact paths, and DOT
rendering. A separate controller hash covers each variant's settings and
generated query-local plans, so controller differences cannot masquerade as
different DFA inputs.
The paired manifest also registers immutable identifier-set digests for all
1,228 achievement cases, all 1,228 temporal cases, the 868 classical reference
cases, and the 360 numeric reference cases. Final table generation recomputes
each digest per method and seed, rejects duplicates and shared omissions, binds
each Raw MOOSE result to the exact corresponding seed model-batch manifest, and
requires all downstream summaries to share one clean source commit. It checks
the registered six-worker, 30-minute, 64-MiB Java-stack, and 8-GiB external-
planner budgets together with pinned tool hashes. The challenge input must
contain all 13 unique registered cases successfully; otherwise no comparison
table is emitted.
Per-instance references receive 30 minutes and 8~GB per problem. For the direct
temporal reference, FOND4LTLf compilation and LAMA search share that single
30-minute deadline rather than receiving separate budgets. We report
paired coverage differences, PAR-2 for timeouts, and plan length only on jointly
solved instances. Tables use short method names and descriptive column phrases;
stable command-line identifiers remain only in machine-readable manifests.

A nonempty execution succeeds only if Jason emits its explicit success marker,
neutral-goal VAL accepts the complete primitive PDDL trace, and replayed states
are accepted by both the gold and predicted DFAs. Failed prefixes are
diagnostics, never plans. For PDDL achievement-goal experiments, VAL instead
checks the original PDDL goal. Rejection is a first-class result, partitioned
into evidence validation, closure, progress, resource, preservation, unsupported
DFA structure, execution failure, timeout, VAL failure, and DFA rejection.

\paragraph{Metrics.}
Atomic metrics are producible-predicate coverage, module closure, held-out
Jason+VAL coverage, selected branches, context literals, primitive and subgoal
steps, ASL bytes, and compilation time. Temporal metrics are controller
acceptance/rejection, Jason success, VAL validity, gold-DFA acceptance, action
count, PAR-2 runtime, append time, controller plans, and maximum trigger fan-out.
For each paired seed and domain, producible-predicate coverage uses one common
denominator: all predicate symbols occurring in positive PDDL action effects.
Each variant is scored by the module triggers it actually emits; missing
closure metadata cannot reduce its denominator.
A separate challenge suite measures false acceptance and false rejection for
unbound variables, non-decreasing recursion, unreleased resource debt, cyclic
threats, forbidden completion additions, and numeric overshoot. Predicate and
action renaming, compatible parameter permutation, object renaming, and
irrelevant-fluent injection test symbol invariance.

\paragraph{Amortization.}
Reusable-library cost is reported as
$T_{E}+T_{C}+k(T_{A}+T_{X})$, where evidence synthesis $T_E$ and domain
compilation $T_C$ occur once, and query append $T_A$ and execution $T_X$ occur
for each of $k$ queries. Per-instance planners incur their complete planning
cost for every query. We report the measured break-even point when one exists;
we do not claim an amortized speedup from incomparable cost components.

\paragraph{Pinned descriptive result.}
Before the registered comparisons finish, the only quantitative execution
claim is the existing hash-locked full-system run. The model is called once for
each of \TranslationInputCount{} deduplicated natural-language inputs. All
\TranslationJSONValidCount{} predictions satisfy the JSON/PDDL contract, and
exact reachable-product exploration finds all \TranslationEquivalentCount{}
predicted DFAs language-equivalent to their gold DFAs. Those predictions expand
to \TEGProblemCount{} bound controllers. All \TEGProblemCount{} execute in
Jason, pass neutral-goal VAL, and are accepted by both DFAs. These values
describe one complete configuration and do not fill any baseline or ablation
cell in Table~\ref{tab:registered-ablations}.

\input{sections/result_domain_table}
\input{sections/result_profile_table}

\paragraph{RQ3--RQ5: temporal execution.}
Table~\ref{tab:profile-results} separates the five declared profiles. Every one
of the \TEGProblemCount{} bound queries compiles and succeeds in Jason; all
\TEGVALSuccessCount{} nonempty traces pass VAL, and both DFAs accept all
\TEGGoldDFAAcceptCount{} traces. No translation/schema error, unsupported DFA,
certificate rejection, Jason failure, timeout, VAL failure, or DFA rejection
occurs in this pinned matrix. The median trace has \TEGMedianActionCount{}
actions, median case runtime is \TEGMedianRuntimeSeconds{} seconds, and
12-worker wall time is \TEGWallRuntimeMinutes{} minutes. Complete coverage
supports the released formulas, bindings, and library hashes only; it is not an
action-strategy completeness theorem for arbitrary PDDL--\ltlf products.

\paragraph{Repetition boundary.}
The atomic and temporal experiments answer different questions. The registered
five-seed matrix estimates variation in evidence discovery and downstream
atomic-library coverage; it is not yet available as a clean result matrix. The
recent temporal experiment instead holds its exact seed-0 atomic libraries
fixed and tests translation, query compilation, execution, and trace semantics
conditional on those SHA-256-identified inputs. Those libraries were generated
with six internal MOOSE workers, so the temporal result remains a valid pinned
conditional experiment but is not evidence for cross-seed robustness or
single-worker synthesis time. Table~\ref{tab:domain-results} reports their
structure only; the two experiments are neither pooled nor presented as the
same estimator.

\paragraph{Independent semantic conformance.}
A separate versioned suite checks hand-labelled positive and negative traces
with both a direct recursive finite-trace evaluator and the real
\texttt{ltlf2dfa}/MONA DFA. Clean run \texttt{\ConformanceRunID} passes
\ConformanceSemanticSuccessCount{}/\ConformanceSemanticCaseCount{} formula
cases and \ConformanceZeroActionSuccessCount{}/\ConformanceZeroActionCaseCount{}
initial-state integration cases. A zero-action execution denotes the singleton
trace $\langle s_0\rangle$ and commits an empty action file. VAL 1.4 cannot
parse an empty plan, so these cases report VAL as not applicable and action
legality as vacuous after PDDL initial-state replay; no noop is inserted. They
remain outside the \TEGProblemCount{} nonempty-plan denominator.
