\section{Certified Atomic-Module Compilation}
\label{sec:method}

\subsection{Method Overview}

GP2PL first maps
$D,\mathcal{I}_{\mathrm{train}},\mathcal E_{\mathrm{raw}}
\longrightarrow\mathcal{M}_D$,
compiling the query-independent certified atomic module core from a PDDL domain,
training split, and generalized-planning evidence. Temporal queries
do not enter this optimization; every later query reuses $\mathcal M_D$.
Algorithm~\ref{alg:atomic} summarizes the procedure.

\begin{algorithm}[htbp]
\caption{Compile the Certified Lifted Atomic Module Core}
\label{alg:atomic}
\begin{algorithmic}[1]
\REQUIRE PDDL domain $D$, training split $\mathcal I_{\mathrm{train}}$,
raw provider artifact $\mathcal E_{\mathrm{raw}}$
\ENSURE Certified lifted atomic module core $\mathcal M_D$ or a rejection diagnostic
\STATE $E_0\leftarrow\textsc{NormalizeEvidence}(\mathcal E_{\mathrm{raw}},D,\mathcal I_{\mathrm{train}})$
\STATE $E\leftarrow\textsc{CanonicalLift}(E_0,D)$
\STATE $T_D(E)\leftarrow\textsc{ProducibleTargets}(E,D)$
\STATE $\mathcal C_E\leftarrow\textsc{ValidateEvidenceMacros}(E,D)$
\STATE $\mathcal C_D(E)\leftarrow
\textsc{LiftSchemas}(D,\mathfrak G,T_D(E),T_D^{\mathbb Z}(E))$
\STATE $\mathcal C^{\checkmark}_{D,E}\leftarrow
\textsc{Certify}(\mathcal C_E\cup\mathcal C_D(E),D)$
\STATE $\mathfrak F_{D,E}\leftarrow
\textsc{FeasibleSubsets}(\mathcal C^{\checkmark}_{D,E},E)$
\IF{$\mathfrak F_{D,E}=\varnothing$}
  \STATE \textbf{reject}
\ENDIF
\RETURN $\mathcal M_D\leftarrow
\textsc{ClingoLexMin}(\mathfrak F_{D,E})$
\end{algorithmic}
\end{algorithm}

Certification and selection operate on trigger--context--body rules before
surface rendering. AgentSpeak(L) rendering preserves the selected triggers,
contexts, bodies, and internal-call relation in $\mathcal M_D$. These selected
rules define the compiler output; AgentSpeak(L) and Jason provide its evaluated
realization. Table~\ref{tab:supported-fragment} states the acceptance and
rejection boundary used by the formal claims and experiments.

\begin{table}[htbp]
\centering
\small
\setlength{\tabcolsep}{2pt}
\begin{tabular}{@{}
>{\raggedright\arraybackslash}p{0.22\columnwidth}
>{\raggedright\arraybackslash}p{0.31\columnwidth}
>{\raggedright\arraybackslash}p{0.39\columnwidth}@{}}
\toprule
Construct & Accepted strategy & Rejection boundary \\
\midrule
Positive predicate & Certified atomic module & No executable achieving branch \\
Numeric equality & Bounded constant-integer progress & Unsupported arithmetic or
incomplete numeric effect \\
Positive conjunction & Threat-safe serialization & Incomplete effects or cyclic
threats \\
Negative predicate & Observe absence or use an exact certified deleter & Synthetic
negated achievement or possible re-addition \\
Mixed Boolean and numeric guard & Complete net-effect and preservation certificate &
Incomplete mixed-effect certificate \\
Boolean alternatives & Separate conjunctive MONA cubes & Explicit disjunction in
the input or one cube; uncertified strategy choice \\
Primitive-state safety & DFA monitor after every primitive action & Completion-only
observation \\
\bottomrule
\end{tabular}
\caption{Supported compilation fragment and fail-closed boundary. Rejection
means that GP2PL emits a structured diagnostic rather than an uncertified
controller or synthetic atomic goal.}
\label{tab:supported-fragment}
\end{table}

\subsection{Evidence Normalization and Canonical Lifting}

An evidence provider is admissible when its raw artifact
$\mathcal E_{\mathrm{raw}}$ can be parsed into a provider-normalized
singleton-goal evidence program $E_0$. Canonical lifting then maps $E_0$ to
$E=\operatorname{Lift}_D(E_0)$, whose rules have the form
\[
e=\langle \bar X_e,C_e^s,g_e,\bar a_e,\nu_e,\pi_e\rangle,
\]
where $C_e^s$ is a partial state condition, $g_e$ is one goal atom,
$\bar a_e$ is a PDDL action macro, $\nu_e$ records numeric conditions and
effects, and $\pi_e$ records provenance. For example, evidence for
\texttt{at(P,L)} may contain a
\texttt{load; move; unload} macro. Admissibility requires every symbol and arity
to occur in $D$, capture-avoiding renaming of local variables, and a singleton
positive achievement target. The lifting map alpha-normalizes provider-local
variables into typed canonical variables, preserves repeated-term sharing, and
leaves constants declared by $D$ fixed. It does not infer a first-order policy
from ground plans: that operation belongs to the external evidence provider.
For example, a provider rule whose declared variables occur as
\texttt{on(block0,block1)} and \texttt{stack(block0,block1)} becomes
\texttt{on(X,Y)} and \texttt{stack(X,Y)}, while a PDDL domain constant remains
unchanged. In the experiments, $E_0$ is obtained from MOOSE readable policies
\cite{Chen2026Moose}. Negative achievement evidence is rejected because
those policies provide no validated action realization for achieving a negated
literal.

\subsection{Action-Schema Candidate Generation}

Let $\operatorname{goalPred}(E)$ contain the predicate symbols targeted by
positive predicate evidence, and let
\[
\operatorname{prod}(D)=
\{p\mid \exists a\in\mathcal A,\ \exists \ell\in\operatorname{add}(a):
\operatorname{pred}(\ell)=p\}.
\]
The domain-wide producible target universe is
\[
T_D(E)=\operatorname{goalPred}(E)\cup\operatorname{prod}(D).
\]
Thus every predicate with a positive producer receives candidate atomic modules
even when the evidence provider never observed it as a goal. A static predicate
appears in no add or delete effect and remains context-only. A dynamic predicate
with no positive producer is not invented as a positive achievement target
unless admissible evidence explicitly supplies a validated realization.
Supported numeric equalities form $T_D^{\mathbb Z}(E)$, the set of admitted
singleton integer-equality targets in $E$; numeric-only evidence does not
suppress $\operatorname{prod}(D)$.

The finite candidate-construction grammar is
\[
\begin{aligned}
\mathfrak G=\{&\text{producer},\text{regression},\text{recursion},
\text{precondition repair},\\
&\text{resource discharge},\text{numeric progress}\}.
\end{aligned}
\]
For the Boolean and numeric targets,
$\textsc{LiftSchemas}$ computes
\[
\mathcal C_D(E)=
\operatorname{Inst}_D(\mathfrak G,T_D(E),T_D^{\mathbb Z}(E)).
\]
It instantiates these constructors using only typed schemas and effects from
$D$ and alpha-normalizes every resulting variable-level branch. Its output is
finite because $D$ and the target signatures are finite, constructor states are
identified up to alpha-normalization, and no requirement/producer state or
resource mode may repeat. For a Boolean target predicate $p$, a producer is
an action schema $a\in\mathcal A$ for which some
$\ell\in\operatorname{add}(a)$ satisfies $\operatorname{pred}(\ell)=p$.
For target set $T$, define the reachable producer--precondition graph
$\mathcal G_D^{\mathrm{prod}}(T)=(V_T,E_T)$ over predicate signatures: an edge
$p\rightarrow r$ belongs to $E_T$ exactly when a producer for $p$ has a positive
dynamic precondition with predicate $r$, and $V_T$ contains the signatures
reachable from $T$ through these edges.
Backward STRIPS regression
\cite{Fikes1972GeneralizedRobotPlans,Alcazar2013Regression,Chen2026Moose}
replaces one
open positive precondition by a producer's preconditions and rejects a step
whose deletes contradict another open requirement. Producer preconditions are
first unified with compatible open requirements, after which only still-unbound
parameters receive fresh variables. Search is enabled only when
$\mathcal G_D^{\mathrm{prod}}(T)$ is acyclic.

Termination does not use an action-depth constant: a path cannot repeat an
alpha-normalized requirement/producer step or resource mode, and only the
shortest body is retained for an equivalent open-requirement and completion-
effect summary. Every candidate is then verified by symbolic progression.
Evidence macros are not truncated. Let $\mathcal C_E$ denote validated evidence
branches and $\mathcal C_D(E)$ the branches generated from $D$ for these target
sets. Cyclic dependencies require provider
evidence or a separately certified recursive module. The resulting candidate
set $\mathcal C_{D,E}=\mathcal C_E\cup\mathcal C_D(E)$ is finite, but
$\mathfrak G$ is not a complete hypothesis language for PDDL planning. Types constrain
unification; an \asl{} realization may expose static sort membership in
contexts, but never as an achievement call or exported action.

\subsection{Candidate Soundness Certificates}

Every candidate carries a machine-checkable certificate. Its conditional
module-completion summary $\Sigma_b$ has Boolean projections
$\mustadd_b$, $\mayadd_b$, and $\maydelete_b$, together with any exact
constant-integer delta and keyed resource-mode transition. Table~\ref{tab:compiler-obligations}
connects each branch constructor to its additional proof obligation and the
failure it excludes.

\begin{table*}[htbp]
\centering
\small
\setlength{\tabcolsep}{3.0pt}
\begin{tabular}{
>{\raggedright\arraybackslash}p{0.17\textwidth}
>{\raggedright\arraybackslash}p{0.42\textwidth}
>{\raggedright\arraybackslash}p{0.32\textwidth}}
\toprule
Branch constructor & Additional acceptance obligation & Excluded failure \\
\midrule
Validated evidence macro & every macro step maps to a PDDL action schema and
symbolic execution establishes the singleton evidence target & invalid or
misbound provider evidence \\
Direct producer & one action adds the target under the certified context and
preserves its completion summary & action chosen only by predicate name \\
Acyclic regression & each open positive requirement is discharged without a
repeated normalized producer role or conflicting delete effect & arbitrary
depth cutoffs or cyclic schema search \\
Relational recursion & a non-negative relation feature strictly decreases and
an already-satisfied, direct-producer, or validated-evidence terminal branch
remains available & recursive oscillation \\
Target-preserving resource discharge & a finite non-repeating mode path restores keyed
availability while preserving the target & unreleased reusable resource \\
Ranked precondition repair & Clingo selects an acyclic call graph with strictly
decreasing natural-number dependency ranks & mutually recursive preparation
without a progress proof \\
\bottomrule
\end{tabular}
\caption{Candidate constructors and soundness obligations for atomic BDI
modules. Every selected
branch additionally satisfies typed binding, symbolic PDDL executability, and
target achievement on return; the resulting core satisfies
$\operatorname{Closed}_D(\mathcal M_D)$.}
\label{tab:compiler-obligations}
\end{table*}

\paragraph{Binding, symbolic execution, and internal-call closure.}
Action and subgoal variables must be bound by the trigger, positive context, or
certified static sort membership; negative contexts and comparisons only
filter. Symbolic progression proves each precondition and final trigger
achievement, recording $\mustadd$, $\mayadd$, and $\maydelete$. For any BDI
plan set $\mathcal R$, define
\[
\operatorname{Closed}_D(\mathcal R)
\Longleftrightarrow
\begin{aligned}[t]
&\forall b\in\mathcal R,\ \forall !q(\bar Y)\text{ occurring in }\beta_b,\\
&\exists b'\in\mathcal R\text{ such that}\\
&\qquad g_{b'}\text{ type-unifies with }q(\bar Y)\\
&\qquad\text{under the type hierarchy of }D.
\end{aligned}
\]
The compiler requires $\operatorname{Closed}_D(\mathcal M_D)$, so every
internal achievement call resolves to a type-compatible selected plan,
including calls reached transitively. This set-level property is a Clingo
obligation; it is distinct from constructing $T_D(E)$. It proves resolution by
signature and type, not applicability of some branch in every reachable state.

\paragraph{Well-founded recursive progress.}
A same-predicate recursive branch $b$ requires a ranking function
$\rho_b:\operatorname{States}(D)\rightarrow\mathbb N$ defined by a non-negative relation count or
anchored acyclic-cone count that strictly decreases and has an already-satisfied,
direct-producer, or validated-evidence terminal branch.
Threats are checked over the preparation graph selected by Clingo, not over
unselected candidates. A branch may add the ranked relation outside the
protected cone only when its trigger preserves the same anchor and a guard proves
that the new anchor differs from the protected one. The ranking feature is used
only in certification and never becomes an agent belief. The construction
follows relational progress witnesses in general-policy work
\cite{Frances2021LearningGeneralPolicies}, but claims only schema-certified
features from the generated candidate set.

\paragraph{Target-preserving resource discharge.}
Effects, not predicate names, identify temporary occupation of a reusable
resource. Suppose an acquisition action deletes an availability fact
$\mathit{free}(R)$ and adds an occupancy relation $\mathit{held}(R,U)$. These
names are illustrative: the compiler infers the shared resource argument $R$
and occupant argument $U$ from the effect structure and obtains the same result
after consistent symbol renaming. The resulting finite abstract mode graph is
$\mathcal G_b^{\mathrm{res}}=(V_b^{\mathrm{res}},E_b^{\mathrm{res}})$: vertices
are alpha-normalized keyed occupancy modes, and an edge labelled by action
schema $a$ records a symbolically executable transition between two modes while
preserving the target. A discharge body is accepted only if symbolic
execution removes the current occupancy mode, restores the corresponding
availability fact, leaves the atomic target true, and does not recreate an
earlier normalized mode. Excluding repeated modes makes the discharge search
finite without assuming that release is one action. Disequalities arise only
when unifying two roles would make a PDDL precondition false.

\paragraph{Ranked cross-module precondition repair.}
Preparing one of several dynamic preconditions may temporarily change another,
such as an object's location. A repair branch may defer an unrelated sibling only
through a schema-inferred single-valued fluent when every deferred sibling has
one producer, all static feasibility witnesses remain, local variables are
alpha-renamed, and the original target is retried before primitive production.
Cross-predicate repair candidates are optional. For a candidate core
$\mathcal M$, Clingo assigns each trigger signature in its selected call graph a
dependency rank $\kappa_{\mathcal M}$ and requires
$\kappa_{\mathcal M}(g)>\kappa_{\mathcal M}(h)$ for every selected call edge
$g\rightarrow h$.
Same-predicate recursion instead requires the
relational certificate above.

\subsection{Feasible-Core Optimization}

Validated evidence macros and certified action-schema-derived candidates enter
one answer-set optimization problem solved with Clingo
\cite{Gebser2018Clingo}. Evidence
rules induce coverage obligations. A selected candidate may cover an evidence
obligation only through alpha-equivalent triggers and bodies, a weaker
conjunctive context with the same body, or an action-schema producer refinement
with the same $\mustadd$, no additional $\maydelete$, and equivalent numeric
and keyed-resource summaries. We write this evidence-coverage relation as
$\operatorname{covers}_D(b,e)$. Syntactic body-prefix similarity is not treated as
semantic equivalence.

The solver also treats every nonrecursive action-schema-derived branch as a
schema-achievement obligation: at least one selected branch must realize the
same certified achievement contract. Let $\mathcal C_D^{\mathrm{nr}}(E)$ denote
these nonrecursive branches and let $\operatorname{realizes}_D(b,c)$ hold when selected
branch $b$ is equivalent to, soundly refines, or gives a certified
target-preserving resource-discharge alternative to schema branch $c$. It
further enforces internal-call closure,
acyclic ranked cross-predicate preparation, and compatibility of selected
effects with recursive rankings. Define
\begin{align*}
\mathcal C^{\checkmark}_{D,E}
  &=\{b\in\mathcal C_{D,E}\mid\operatorname{Cert}_D(b)\},\\
\mathfrak F_{D,E}
  &=\{\mathcal M\subseteq\mathcal C^{\checkmark}_{D,E}\mid
      \mathcal M\models\Omega_{D,E}\land
      \operatorname{Closed}_D(\mathcal M)\}.
\end{align*}
where $\Omega_{D,E}$ contains evidence coverage, schema-achievement coverage,
preparation acyclicity, and recursive-ranking compatibility. Resource discharge
is already part of the per-branch predicate
$\operatorname{Cert}_D$. In particular, schema-achievement coverage requires
\[
\forall c\in\mathcal C_D^{\mathrm{nr}}(E),\quad
\exists b\in\mathcal M:\operatorname{realizes}_D(b,c).
\]
The optimizer uses the
lexicographic objective
\begin{multline*}
\mathcal M_D=\operatorname*{lexargmin}_{\mathcal M\in\mathfrak F_{D,E}}
\bigl\langle-N_{\mathrm{rel}}(\mathcal M),-N_{\mathrm{prep}}(\mathcal M),
|\mathcal M|,\\
N_C(\mathcal M),N_\beta(\mathcal M)\bigr\rangle,
\end{multline*}
where $N_{\mathrm{rel}}$ and $N_{\mathrm{prep}}$ count compatible
relation-ranked recursion and acyclic precondition-repair capabilities,
$N_C$ counts context literals, and $N_\beta$ counts body steps. Clingo computes
this core directly for the grounded finite instance. The selected
trigger--context--body rules, rather than their AgentSpeak rendering, define
$\mathcal M_D$. The claim is global only within
$\mathcal C^{\checkmark}_{D,E}$, not over ungenerated programs.

The compiler rejects an untyped predicate with several producers when their
safe use requires different nested precondition-repair branch sets. For
example, if one
\texttt{at/2} producer preserves a ranked relation and another may increase it,
static role names alone do not prove that their object sets are disjoint. The
compiler neither recognizes those names nor assumes disjointness.

\section{DFA-Guided Composition of Temporally Extended Goals}
\label{sec:temporal-composition}

\paragraph{Validated parametric input.}
GP2PL does not prescribe a controlled surface grammar for user requests. In the
evaluated front end, a source utterance is accompanied by the public PDDL symbol
catalogue, declared typed parameters, and binding constraints. Translation
produces a structured specification
$\tau_q=\langle\iota_q,\varphi_q,\mu_q,\Theta_q,\Gamma_q\rangle$, where
$\iota_q$ is its identifier, $\varphi_q$ is an \ltlf{} formula over proposition
symbols, $\mu_q$ maps each proposition to a lifted PDDL predicate or integer
equality, $\Theta_q$ is the typed parameter signature, and $\Gamma_q$ contains
binding constraints. Before automaton construction, validation requires a
bijective formula-to-atom mapping, membership in the domain vocabulary, and
correct arities, types, numeric values, temporal operators, and literal-only
negation. Invalid
specifications are rejected rather than repaired. The controlled utterances in
the released benchmark define the evaluation distribution, not a required
user-facing grammar; the serialized schema and translation protocol appear in
the Technical Supplement, Sec.~4.2.

A runtime binding is a type-consistent map
$\theta_q:\bar X_q\rightarrow O_I$ satisfying the type signature
$\Theta_q:\bar X_q\rightarrow\mathcal T$ and constraints $\Gamma_q$; the validated bound
query is $\widehat\tau_q=(\tau_q,\theta_q)$, and $g_q$ denotes its generated
top-level BDI goal. Applying $\theta_q$ substitutes declared parameters and
leaves PDDL domain constants fixed. For $a\in AP_q$, its bound PDDL-state
valuation is
\[
a\in\mathit{val}_q(s)
\Longleftrightarrow
\begin{cases}
p(\bar t)\theta_q\in s_{\mathcal P},
  & \mu_q(a)=p(\bar t),\\
s_{\mathcal F}(f(\bar t)\theta_q)=c,
  & \mu_q(a)=[f(\bar t)=c].
\end{cases}
\]
Thus predicate propositions denote grounded Boolean facts and numeric
propositions denote exact bounded-integer equalities in the same observed
state. As defined in Sec.~\ref{sec:background}, LTLf2DFA/MONA produces
\[
\mathcal D_q=
\langle Q_q^{\mathrm{dfa}},2^{AP_q},\delta_q,q_q^0,F_q\rangle
\]
and its guarded transitions
\cite{Fuggitti2020LTLf2DFA,Henriksen1995MONA}. GP2PL attaches a controller to
each progress edge $(q_s,\chi,q_t)$ with
$d_q(q_t)<d_q(q_s)<\infty$; an accepting \texttt{true} self-loop needs none.

Same-source/same-target valuation cubes may share a common achievement
objective, but the monitor still evaluates every complete cube. They remain
separate conjunctive alternatives in MONA's transition relation, not one
admitted disjunctive guard.

Each execution applies $\theta_q$: $\mathit{on}(X,Y)$ under
$X{=}b_1,Y{=}b_4$ calls $!\mathit{on}(b_1,b_4)$ without grounding the shared
$+!\mathit{on}(X,Y)$ module. A binding inconsistent with the validated query
specification is rejected.
For guard
$\chi=\bigwedge_{G\in P_\chi^+}G\land
\bigwedge_{N\in P_\chi^-}\neg N$, the signed transition obligation is
$\mathcal O_\chi=\langle P_\chi^+,P_\chi^-\rangle$. Compilation computes
conditional module-completion summaries and a threat-induced precedence graph, building
on BDI goal-interference analyses that
distinguish definite from potential plan effects
\cite{Thangarajah2003GoalInterference}. An acyclic graph is topologically
ordered; a cyclic graph requires either an occurrence-specific preservation
portfolio or a joint guard-establishment branch $b_\chi^{\mathrm{joint}}$ whose
complete net effect establishes the entire guard. The transition is rejected if neither
can be certified.
When $b_\chi^{\mathrm{joint}}$ establishes the complete guard, it is available
from every positive occurrence it establishes and retains every query variable
appearing in $\mathcal O_\chi$. Thus an already-true first literal cannot hide
the branch from an unmet sibling or discard a query binding.
The certified serialization $\boldsymbol\ell_\chi$ is rendered as a balanced
binary transition-repair tree by
Algorithm~\ref{alg:temporal}.

\subsection{Conditional Module-Completion Summaries}

A relational fixed point over selected calls computes $\Sigma_b$, the effects
observable when branch $b$ returns. Query arguments remain anchors; local
variables are alpha-normalized and typed; effects retain branch conditions.
Sequential composition records net
polarity, so an internally deleted-and-restored atom is not a completion delete.

For occurrence $G_i\in P_\chi^+$, a preservation portfolio
$\Pi_{\chi,i}\subseteq\mathcal M_D$ is the set of certified branches permitted
to establish $G_i$ under the obligations protected at that position. The
threat-induced precedence relation is defined by
$G_j\prec_\chi G_i$ exactly when a feasible $\maydelete$ effect of a branch in
$\Pi_{\chi,j}$ unifies with $G_i$. The graph
$\mathcal G_\chi=(P_\chi^+,\prec_\chi)$ therefore requires $G_j$ to be repaired
before $G_i$.
Single-valued invariants reject conjunctions requesting two values for one key.

A negative guard $\neg N$ never becomes $!\neg N$. Positive repairs must exclude
feasible $\mayadd(N)$. If $N$ holds, a signed leaf may call only a finite branch
with exact $\operatorname{MustDelete}(N)$ that establishes a positive sibling and
preserves all obligations. A negative-only edge observes absence. Direct-
deleter parameters bind through compatible sibling add effects before range
restriction, without recognizing fluent names.

\subsection{Preservation Portfolios for Cyclic Interference}

For a cycle, each $\Pi_{\chi,i}$ retains only branches whose exact completion
preserves prior positives and negatives. In the AgentSpeak realization, a
query-local trigger exposes exactly the bodies rendered from that branch set.
Portfolios are certified per ordered literal occurrence: two occurrences of
$p$ may need different portfolios when
their protected prefixes differ, while identical portfolios may share one
restricted module in the AgentSpeak realization.
Contexts remain alternative applicability cases, not simultaneous obligations.
A recursive $!p;!g$ branch also requires a complete preserving summary for $p$
and redirects $!g$ to the occurrence's restricted module. Certificates record
source branches by literal index and claim recursive preservation coverage only
when at least one branch was selected. For a binary support relation $R$, the
same-predicate rank $\rho_b$ may be instantiated as support depth $\rho_R(x)$
only when its grounded request graph is
functional and acyclic in every reachable state. Every selected branch must
still preserve earlier achievements; otherwise compilation rejects.

\subsection{Numeric and Joint Guard-Establishment Certificates}

A positive integer equality may call a complete constant-delta action: unit
deltas require strict direction, larger deltas the exact predecessor. Mixed
guards use complete action-only net effects. For example, a body for
$\mathit{at}(P,L)\land \mathit{fuel}(V)=3$ must symbolically establish
$\mathit{at}(P,L)$ and the exact integer value 3 in its final state while
preserving every other signed guard obligation. A cyclic guard admits one
joint guard-establishment action only if the same proof covers the complete
successor valuation required by $\mathcal O_\chi$. Negated equality is
observation-only without a certified change-away action; arbitrary arithmetic
and uncertified sequential composition are outside the supported fragment.

Numeric preparation must strictly reduce the lexicographic ranking
$\rho_{\mathrm{num}}=\langle n_{\mathrm{miss}},d_{\mathrm{target}}\rangle$,
where $n_{\mathrm{miss}}$ is the number of missing producer preconditions and
$d_{\mathrm{target}}(s)=|f(s)-c|$ is the bounded deficit for target equality
$f=c$. It must preserve
other conditions and leave the target fluent unchanged until the establishing
step. Compilation rejects the
candidate when no decreasing tuple can be certified.

For strong Until, let $\mathcal W_{q_s}$ be the set of non-progress self-loop
guards at source $q_s$. Their signed source invariants are
\[
\mathcal I_{q_s}=\langle I_{q_s}^+,I_{q_s}^-\rangle,
\quad I_{q_s}^{\pm}=\bigcap_{\chi\in\mathcal W_{q_s}}P_\chi^{\pm}.
\]
Before the single progress literal, no action prefix may delete a positive
invariant or add a negative one; the establishing step may consume the left
operand. Repeated unit progress separates action and protected objects by a
derived non-unification guard, and unavoidable consumption requires the exact
predecessor. An observed target equality is the base case.

Inserted auxiliary variables are alpha-renamed away from query and producer
variables, preventing textual collisions from imposing unrelated constraints.

\subsection{Balanced Binary Transition-Repair Trees}

Let $\boldsymbol\ell_\chi=(\ell_1,\ldots,\ell_n)$ be a certified serialization
of $\mathcal O_\chi$ and let $\Pi_{\chi,i}$ be the optional preservation
portfolio for occurrence $\ell_i$. A positive leaf
observes or achieves its atom; a negative leaf observes absence or invokes its
exact certified deleter. The tree indexes these occurrences, not DFA states or
primitive actions. For query $q$, let $\mathcal R_{q_s,\chi}$ denote the plans
for one source-state transition obligation and let the query-local plan set $\mathcal Q_q$
contain the top-level dispatcher together with every accepted
$\mathcal R_{q_s,\chi}$. Algorithm~\ref{alg:temporal} constructs one such
transition plan set.

\begin{algorithm}[htbp]
\caption{Compile a Certified DFA Progress Controller}
\label{alg:temporal}
\begin{algorithmic}[1]
\REQUIRE DFA source $q_s$, guard $\chi$, signed obligation $\mathcal O_\chi$,
selected atomic core $\mathcal M_D$
\ENSURE Transition-repair plan set $\mathcal R_{q_s,\chi}$ or rejection
\STATE $\boldsymbol\Sigma\leftarrow
\{\Sigma_b\mid b\in\mathcal M_D\}$
\STATE $(\boldsymbol\ell_\chi,\boldsymbol\Pi_\chi)\leftarrow
\textsc{CertifySerialization}(\mathcal O_\chi,\boldsymbol\Sigma)$
\IF{$(\boldsymbol\ell_\chi,\boldsymbol\Pi_\chi)=\bot$}
  \STATE \textbf{reject}
\ENDIF
\STATE emit entry $r_{q_s,\chi}\leftarrow !R_\chi[1,n];!c_{q_s,\chi}$
\STATE \textbf{procedure} $\textsc{Build}(i,j)$
\IF{$i=j$}
  \STATE emit the satisfied plan for signed literal $\ell_i$
  \IF{$\Pi_{\chi,i}\ne\varnothing$}
    \STATE emit the unmet plan restricted to $\Pi_{\chi,i}$
  \ENDIF
\ELSE
  \STATE $m\leftarrow\lfloor(i+j)/2\rfloor$
  \STATE emit $R_\chi[i,j]\leftarrow !R_\chi[i,m];!R_\chi[m+1,j]$
  \STATE $\textsc{Build}(i,m);\ \textsc{Build}(m+1,j)$
\ENDIF
\STATE \textbf{end procedure}; $\textsc{Build}(1,n)$
\STATE emit completion test $c_{q_s,\chi}$: return if monitor $\ne q_s$;
otherwise call $!r_{q_s,\chi}$
\RETURN $\mathcal R_{q_s,\chi}$
\end{algorithmic}
\end{algorithm}

A complete transition-repair pass is
$\operatorname{Pass}_{q_s,\chi}=R_\chi[1,n]\cdot c_{q_s,\chi}$, where $\cdot$
denotes sequential controller composition. The monitor still advances after
every primitive action, but
the controller tests source-state exit only after the whole pass. It does not
interrupt an atomic call or short-circuit the remaining ranges.

Suppose the monitor leaves $q_s$ while a branch in $\Pi_{\chi,i}$ executes.
After that call
returns, every suffix leaf either observes its literal without acting or selects
only a branch from its already certified $\Pi_{\chi,j}$. Applicability is rechecked at invocation;
an applicable call is PDDL-executable and preserves the established signed
prefix at module return, while a missing call fails without inserting an
uncertified action. Every primitive suffix action is observed by the same DFA,
so it may traverse further actual edges, including a rejecting edge, but cannot
manufacture a transition or report false acceptance. Only after the suffix
finishes does the completion test return to top-level dispatch at the current state or replay
when that state is still $q_s$. Consequently, the controller cannot report an
accepting transition that the monitor did not take. The suffix need not be a
complete strategy for a newly reached state.

For $n=1$, $R_\chi[1,1]$ is one leaf followed by the same completion test. Balancing
does not choose the certified order or reduce primitive action count; it bounds
trigger fan-out by two and nesting depth logarithmically while retaining linear
work per pass. The repeated state check is an instance of the declarative goal
plan pattern \cite{Hubner2006DeclarativeGoals}; the balanced dispatch structure
is GP2PL's controller realization.

\subsection{Maintained Domain Library}

The atomic compiler constructs $\mathcal M_D$ once. Temporal compilation for a
query $q_i$ unions its dispatcher and transition plan sets into
$\mathcal Q_{q_i}$. For an ordered sequence $q_1,\ldots,q_k$, the sole
maintained library is
\[
\mathcal L_D^{[k]}=
\mathcal M_D\cup\bigcup_{i=1}^{k}\mathcal Q_{q_i},
\qquad
\mathcal L_D^{[0]}=\mathcal M_D.
\]
The \asl{} realization serializes $\mathcal L_D^{[k]}$ as one per-domain plan
library; $\mathcal M_D$ is its stable query-independent subset, not a second
library. Appending $\mathcal Q_{q_i}$ neither reconstructs nor reoptimizes the
atomic core. Query-local goals are control symbols, not PDDL fluents or exported
actions. Jason records the complete primitive trace, which VAL checks against
the PDDL problem \cite{Howey2004VAL}.

If the initial valuation is accepting, the committed plan is empty and the
trace is $\langle s_0\rangle$, not an empty trace or inserted noop; a noop could
change strong-Next or Until truth under finite-trace semantics
\cite{DeGiacomoVardi2013}.

\section{Conditional Guarantees}
\label{sec:guarantees}

The guarantees concern accepted artifacts under their stated premises. They are
conditional on the generated certified candidate set and certificate-accepted
temporal subset; they do not assert complete candidate generation, universal
AgentSpeak optimality, or complete strategy synthesis for arbitrary PDDL
domains and DFAs.

\paragraph{Atomic branch soundness.}
\textbf{Proposition 1.} If an emitted $p(\bar X)$ branch's context holds under a
type-consistent grounding and called modules satisfy their certified achievement
and completion summaries, the branch
is executable and returns with $p(\bar X)$. Binding grounds each term; induction
over symbolic progression establishes preconditions and final achievement;
recursion uses its ranking and preparation rechecks the producer context.
\hfill$\square$ The proposition establishes local soundness, not
reachable-state coverage.

\paragraph{Certified-candidate-set optimality.}

\textbf{Proposition 2.} For the grounded selection instance defined above, a
Clingo optimum $\mathcal M_D$ belongs to $\mathfrak F_{D,E}$ and minimizes the
declared lexicographic cost over that feasible family.

The Technical Supplement proves a bidirectional correspondence between answer
sets and feasible subsets. The result does not imply complete candidate
generation or optimality over all AgentSpeak programs.

\paragraph{Supported transition-composition soundness.}

\textbf{Proposition 3.} Let $\mathcal O_\chi=\langle
P_\chi^+,P_\chi^-\rangle$ be an accepted signed transition obligation. If each
required signed repair has an applicable certified branch,
each selected call terminates, and later repairs preserve earlier obligations,
then at the end of a complete transition-repair pass the controller retries
exactly when the monitor reports the source state; otherwise it returns to
dispatch at the
current state. Every intervening state change is an actual DFA edge enabled by
the observed valuation, and a rejecting state cannot report success.

\textbf{Proof sketch.} Induct over the certified order: each leaf establishes
its signed literal; later summaries exclude deleting prior positives or adding
forbidden atoms. If the source is left during a call, the remaining suffix still
contains only these certified calls and observations, so it preserves the prefix
at return boundaries. The monitor evaluates full cubes after every action and
cannot be written by the controller; the suffix can therefore cause only real,
fully observed DFA transitions. After the full pass, the completion plan tests
the current state and either retries or returns to top-level dispatch.
\hfill$\square$ This does not imply complete action-strategy synthesis or
guarantee that a post-exit suffix avoids a rejecting state.

\paragraph{Balanced transition-repair complexity.}

\textbf{Proposition 4.} For $n$ signed literals and $e$ repair plans, the tree
has $2n+e+2$ plans, fan-out two, depth $O(\log n)$, and $O(n)$ visits per pass,
while preserving order. The count follows from $n$ satisfied leaves, $e$ repair
leaves, $n-1$ internal nodes, one entry, and two completion plans; midpoint
splitting and range induction establish depth and order. \hfill$\square$

\paragraph{Monitor trace fidelity.}

\textbf{Proposition 5.} Let
$\xi=s_0,a_1,s_1,\ldots,a_k,s_k$ be the PDDL execution observed by the monitor,
and let $z_0=\delta_q(q_q^0,\mathit{val}_q(s_0))$. After every successful
primitive action $a_i$, the runtime monitor state equals
$z_i=\delta_q(z_{i-1},\mathit{val}_q(s_i))$. Hence leaving a controller source state
is exactly an exit in the deterministic automaton run over the observed finite
trace.

The result follows by induction on primitive actions: the environment first
applies the PDDL successor and then evaluates the deterministic transition
function on the resulting valuation. Controller plans cannot write monitor
beliefs. \hfill$\square$

\paragraph{Initial-acceptance trace semantics.}

\textbf{Proposition 6.} If the initial valuation is accepting, successful
execution commits no primitive action and denotes the one-state finite trace
$\langle s_0\rangle$. It is therefore equivalent to evaluating the query on the
initial state and is not equivalent to inserting an arbitrary noop.

This is immediate from finite-trace semantics and Proposition~5: the initial
valuation is consumed before action execution, while a noop would introduce a
second observed position and could change strong-Next or Until truth.
\hfill$\square$

Complete proofs, including the binding, symbolic-progression,
relational-ranking, resource-discharge, and tree-order lemmas used by these
propositions, appear in the Technical Supplement, Sec.~2.
