\section{Related Work and Positioning}
\label{sec:related-work}

\paragraph{Generalized planning.}
Generalized planning seeks compact policies or programs that solve families of
planning instances
\cite{Jimenez2019ReviewGeneralizedPlanning,Srivastava2011DirectedSearch,
Chen2026Moose}.
Feature approaches learn abstractions or policies from small examples
\cite{Bonet2018Features,Frances2021LearningGeneralPolicies,
Bonet2024GeneralPoliciesWidth}. Policy sketches encode serialized subgoal
structure, with answer-set optimization used to learn compact rules
\cite{Drexler2022LearningSketches,Drexler2024SubgoalSketches}; PG3 uses lifted
decision lists \cite{Yang2022PG3}, and BFGP++ uses terminating programs
\cite{SegoviaAguas2022BFGPpp}. MOOSE learns lifted condition-to-macro rules by
singleton-goal regression \cite{Chen2026Moose}. These methods learn reusable
behavior. Recent BDI integrations instead invoke a classical planner at run
time \cite{Xu2024OperationalSemantics} or use generalized planning directly for
means--ends reasoning \cite{Meneguzzi2025GeneralisedBDIPlanning}. GP2PL instead
addresses the representation-compilation problem of transforming
singleton-goal evidence, together with the domain's producible target universe
and action-schema-derived candidates, into a query-independent atomic module core
$\mathcal M_D$ that can be extended by later queries.

\paragraph{Modular policies and BDI plan libraries.}
Policy-reuse languages allow one generalized policy to invoke another with
parameters \cite{Bonet2024PolicyReuse}, motivating internal calls such as
\texttt{!clear(Y)}. Procedural BDI languages likewise organize behavior as
event-triggered plans selected from a context-dependent library
\cite{MeneguzziDeSilva2015PlanningBDI,DeSilva2020BDIArchitectures}; \asl{}
formalizes these plans and Jason executes an extension
\cite{Rao1996AgentSpeak,Bordini2007Jason}. Plan patterns for declarative goals
recheck intended states after procedural return \cite{Hubner2006DeclarativeGoals};
plan-failure work separates a failed attempt from continued goal pursuit
\cite{Sardina2007PlanFailure}. Definite and potential effect summaries detect
interference between BDI goals \cite{Thangarajah2003GoalInterference}. GP2PL
adopts these execution principles and derives certified modules,
conditional module-completion summaries, and declarative completion controllers from
planning evidence and PDDL schemas.

\paragraph{Finite-trace temporal control.}
\ltlf{} provides finite-trace semantics and admits automata-based reasoning
\cite{DeGiacomoVardi2013}. We instantiate the automata-theoretic translation
with the MONA-backed \texttt{ltlf2dfa} construction
\cite{Fuggitti2020LTLf2DFA,Henriksen1995MONA}. Full temporal synthesis
synthesizes over the domain--automaton product
\cite{DeGiacomoRubin2018AutomataFOND}; GP2PL instead composes certified progress
over an independent BDI library. Temporally extended goals have also been given
explicit agent-language semantics \cite{Hindriks2009TemporalGoals}; we retain
standard plan execution and add a query-local monitor. Plan4Past's fixed-planner
comparison motivates our paired temporal design \cite{Bonassi2023Plan4Past}.
FOND4LTLf \cite{DeGiacomoFuggitti2021FOND4LTLf} is a task-level reference on its
accepted subset, using LAMA as its fixed planner.

\paragraph{Natural-language temporal specifications.}
Prior systems map natural-language instructions to temporal logic through
extensible pattern libraries \cite{FuggittiChakraborti2023NL2LTL}, lifted
language-to-logic translation \cite{Chen2023NL2TL}, or modular grounding in new
environments \cite{Liu2023Lang2LTL}; related planning work extracts ordering and
blocking constraints into solver specifications \cite{Guo2025CaStL}. Prompt
programs can constrain structured outputs \cite{BeurerKellner2023Prompting},
but prompt-based semantic parsers remain sensitive to input perturbations
\cite{Zhuo2023RobustSemanticParsing}. Our evaluated translation protocol
therefore constrains the schema and vocabulary, then validates typed,
externally bound parametric \ltlf{} by exact finite-trace language equivalence
before GP2PL compilation. The complete prompt appears in the Technical
Supplement, Sec.~4.2.

\paragraph{Position of this work.}
Generalized planners target policies or programs, BDI programming assumes a
plan library, and temporal synthesis targets an automaton strategy. GP2PL
connects them through certificate-carrying compilation. The broader
proof-carrying principle attaches machine-checkable evidence to executable
artifacts
\cite{Necula1997ProofCarryingCode}; GP2PL's certificates are narrower symbolic
planning proof obligations. GP2PL contributes the certified construction of a reusable
atomic module core together with preservation-safe query plans in the maintained
BDI library $\mathcal L_D^{[k]}$; goal regression, answer-set solving,
AgentSpeak semantics, and \ltlf-to-DFA translation remain prior components.
