\section{Formal Framework}
\label{app:formal-framework}

This appendix makes explicit the mathematical objects, assumptions, and proofs
summarized in the main paper. The framework is defined at the level of typed
planning transitions \cite{McDermott1998PDDL,FoxLong2003PDDL21} and
trigger--context--body BDI rules
\cite{Rao1996AgentSpeak,MeneguzziDeSilva2015PlanningBDI}. MOOSE, Clingo,
AgentSpeak(L), and Jason instantiate four interfaces; none is built into the
semantic definitions below.

\subsection{Planning and Evidence}

\begin{definition}[Typed planning domain]
A domain is
$D=\langle\mathcal P,\mathcal F,\mathcal A,\mathcal T\rangle$.
$\mathcal P$ is a finite set of typed predicate schemas,
$\mathcal F$ a finite set of integer-valued function schemas,
$\mathcal A$ a finite set of deterministic action schemas, and
$\mathcal T$ a finite type hierarchy. A problem instance is
$I=\langle D,O_I,s_I^0,G_I\rangle$, where $O_I$ is its finite typed object set,
$s_I^0$ its initial state, and $G_I$ its PDDL achievement condition. Write
$\operatorname{States}(D,O_I)$ for the typed state set admitted by $D$ over
$O_I$. A state
$s=(s_{\mathcal P},s_{\mathcal F})$ is a Boolean valuation of ground predicates
and a bounded integer valuation of supported ground functions. An applicable
ground action $a$ has successor $\gamma(s,a)$ given by its PDDL add, delete, and
constant-integer effects.
\end{definition}

\begin{definition}[Singleton-goal evidence]
A raw provider artifact $\mathcal E_{\mathrm{raw}}$ is admissible when it can be
normalized into a finite evidence program $E_0$, understood as a finite rule set.
A normalized evidence rule is
\[
e=\langle\bar X_e,C_e^s,g_e,\bar a_e,\nu_e,\pi_e\rangle.
\]
$\bar X_e$ are typed variables, $C_e^s$ is a partial state condition, $g_e$ is
one positive predicate literal or supported integer equality, $\bar a_e$ is a
finite PDDL action macro, $\nu_e$ records numeric conditions and effects, and
$\pi_e$ records provenance. The program $E_0$ is admissible for $D$ only if
all symbols, arities, types, and macro actions map to $D$, and symbolic execution
under $C_e^s$ establishes $g_e$.
\end{definition}

\begin{definition}[Canonical lifting]
For an admissible provider-normalized program $E_0$, canonical lifting is the
map $\operatorname{Lift}_D(E_0)=E$ that alpha-renames every provider-local
variable to a typed canonical variable. The map preserves repeated-variable sharing
within and across a rule's goal, context, and macro actions; constants declared
by $D$ remain constants. Variables introduced while instantiating typed action
schemas are fresh with respect to the lifted evidence rule. Thus lifting is
invariant to provider-local variable names and cannot replace a training object
by an unrelated constant. It is a representation-level canonicalization of
first-order evidence, not an inference procedure from ground plans.
\end{definition}

MOOSE readable policies \cite{Chen2026Moose} are the evaluated evidence
instantiation. For example,
a readable rule may associate the singleton goal $on(X,Y)$ with context
$holding(X)\land clear(Y)$ and action $stack(X,Y)$. GP2PL does not assume that
this evidence already implements $clear(Y)$ or covers every reachable context.

\begin{definition}[Producible target universe]
For domain $D$, let
$\operatorname{prod}(D)=\{p\mid p$ occurs in a positive add effect of some
$a\in\mathcal A\}$. If $\operatorname{goalPred}(E)$ is the set of positive
predicate symbols targeted by admissible evidence, then
\[
T_D(E)=\operatorname{goalPred}(E)\cup\operatorname{prod}(D).
\]
$T_D(E)$ is the Boolean target universe used for action-schema-derived atomic
candidate generation. Predicates with no add or delete effect are static and may
occur only as contexts. A delete-only dynamic predicate is not an
action-schema-derived
positive target. The numeric target set
$T_D^{\mathbb Z}(E)$ contains the admitted singleton integer equalities in $E$
and does not remove any member of $\operatorname{prod}(D)$.
\end{definition}

For target set $T$, the producer--precondition dependency graph is
$\mathcal G_D^{\mathrm{prod}}(T)=(V_T,E_T)$. Its vertices are predicate
signatures reachable from $T$; $p\rightarrow r$ is in $E_T$ exactly when an
action-schema producer for $p$ has a positive dynamic precondition with
predicate $r$. Action-schema regression is admitted only on an acyclic such
graph. This is a structural condition over renamed schemas, not a domain label.

Throughout, an obligation is a condition the compiler must establish. A witness
is finite evidence for one obligation, such as a symbolic-progression derivation,
decreasing rank, or resource-discharge path. A certificate is the finite,
machine-checkable record that identifies an artifact's applicable obligations
and their witnesses; for example, a transition certificate records portfolios,
effect-summary completeness, threats, and a safe order. Certification checks that record,
and certified denotes only an artifact that passes all required checks.

\begin{table*}[htbp]
\centering
\small
\setlength{\tabcolsep}{3pt}
\begin{tabular}{@{}
>{\raggedright\arraybackslash}p{0.18\textwidth}
>{\raggedright\arraybackslash}p{0.27\textwidth}
>{\raggedright\arraybackslash}p{0.18\textwidth}
>{\raggedright\arraybackslash}p{0.27\textwidth}@{}}
\toprule
\multicolumn{2}{c}{Atomic compilation} &
\multicolumn{2}{c}{Temporal composition and execution} \\
Symbol & Unique semantic role & Symbol & Unique semantic role \\
\midrule
$D,s,\gamma$ & Typed planning domain, world state, and deterministic PDDL
successor function. &
$\tau_q,\theta_q,\widehat\tau_q$ & Unbound specification, external binding,
and validated bound query. \\
$I,O_I,\mathcal I_{\mathrm{train}}$ & Problem instance, its typed object set,
and finite training instance set. &
$\iota_q,g_q$ & Query identifier and generated top-level BDI goal. \\
$\mathcal E_{\mathrm{raw}},E_0,E,e$ & Raw provider artifact, provider-normalized
program, canonically lifted finite evidence program, and one evidence rule. &
$\mu_q,\Theta_q,\Gamma_q$ & Proposition map, parameter type signature, and
binding constraints. \\
$T_D(E),T_D^{\mathbb Z}(E)$ & Boolean producible target universe and admitted
integer-equality targets. &
$\mathcal D_q,\mathit{val}_q,d_q$ & Deterministic finite automaton, PDDL-state
valuation, and acceptance distance. \\
$\operatorname{prod}(D),\operatorname{goalPred}(E)$ & Predicate symbols with
producers and positive predicate evidence. &
$q_s,\chi,P_\chi^+,P_\chi^-$ & Source state, transition guard, and its positive
and negative occurrences. \\
$\mathcal G_D^{\mathrm{prod}}(T)$ & Producer--precondition dependency graph
reachable from targets $T$. &
$\mathcal O_\chi$ & Signed transition obligation induced by $\chi$. \\
$\mathfrak G,\operatorname{Inst}_D$ & Domain-independent candidate-construction
grammar and lifted typed-schema instantiation. &
$\Pi_{\chi,i},\mathcal G_\chi,\prec_\chi$ & Preservation portfolio,
threat-induced precedence graph, and relation. \\
$\mathcal C_E,\mathcal C_D(E),\mathcal C_{D,E}$ & Validated evidence branches,
action-schema-derived branches, and generated union. &
$\boldsymbol\ell_\chi$ & Serialization recorded by the transition certificate. \\
$b,\Sigma_b,\operatorname{Cert}_D(b)$ & One candidate branch, its conditional
completion summary, and branch-certification predicate. &
$\mathcal W_{q_s},\mathcal I_{q_s}$ & Waiting self-loop guards and signed source
invariants. \\
$\rho_b,\kappa_{\mathcal M},\mathcal G_b^{\mathrm{res}}$ & Same-predicate recursive
ranking, dependency ranking, and resource-mode graph. &
$\rho_{\mathrm{num}},b_\chi^{\mathrm{joint}}$ & Numeric-progress ranking and
complete joint guard-establishment branch. \\
$\operatorname{Closed}_D,\operatorname{covers}_D$ & Internal-call closure
predicate and evidence-coverage relation. &
$R_\chi[i,j],c_{q_s,\chi}$ & Transition-repair subtree and source-state
completion test. \\
$\mathcal C_D^{\mathrm{nr}}(E),\operatorname{realizes}_D$ & Nonrecursive
schema obligations and schema-realization relation. &
$\operatorname{Pass}_{q_s,\chi},\mathcal R_{q_s,\chi}$ & Complete repair pass
and plan set for one source-state obligation. \\
$\mathcal C^{\checkmark}_{D,E},\Omega_{D,E},\mathfrak F_{D,E}$ & Branch-certified
candidate set, set-level obligations, and feasible selection family. &
$\mathcal Q_q,\mathcal L_D^{[k]}$ & Query-local controller plans and maintained
library after $k$ appends. \\
$\mathcal M_D$ & Lexicographically optimal feasible atomic module core. &
$\xi,z_i$ & Committed PDDL trace and runtime DFA state after observation $i$. \\
$N_{\mathrm{rel}},N_{\mathrm{prep}},N_C,N_\beta$ & Counts of selected
relation-ranked recursion, acyclic precondition repair, context literals, and
body steps. &
$\Phi_{\mathrm{ach}},\Phi_{\mathrm{syn}}$; \newline
$\Phi_{\mathrm{bench}},\Phi_{\mathrm{cert}}$; \newline
$\operatorname{emb}$ & Atemporal achievement fragment, declared input syntax,
evaluated TEG family, certified subset, and execution embedding. \\
\bottomrule
\end{tabular}
\caption{Canonical GP2PL notation. A symbol is not reused across evidence,
candidate generation, selected libraries, queries, and runtime traces.}
\label{tab:canonical-notation}
\end{table*}
\FloatBarrier

\subsection{Parametric LTLf Syntax and Semantics}

\begin{definition}[Declared input syntax]
For proposition alphabet $AP_q$, the admitted formula language
$\Phi_{\mathrm{syn}}$ is generated by
\[
\begin{aligned}
\ell &::= a\mid\neg a,\\
\varphi &::= \ell\mid(\varphi\land\varphi)\mid F\varphi\mid X\varphi\\
&\phantom{{}::={}}\mid(\varphi U\varphi),\qquad a\in AP_q.
\end{aligned}
\]
Negation is restricted to atoms and is written \texttt{!} in serialized input.
The syntax contains no disjunction, global, release, weak-next, implication,
equivalence, or quantifier. Parsing a formula in $\Phi_{\mathrm{syn}}$ does not
assert that a PDDL action strategy exists.

The atemporal achievement fragment $\Phi_{\mathrm{ach}}\subset
\Phi_{\mathrm{syn}}$ contains exactly the formulae generated by
$\psi::=\ell\mid(\psi\land\psi)$. Formulae in
$\Phi_{\mathrm{syn}}\setminus\Phi_{\mathrm{ach}}$ are temporally extended goals.
The execution embedding is
\[
\operatorname{emb}(\varphi)=
\begin{cases}
F\varphi,&\varphi\in\Phi_{\mathrm{ach}},\\
\varphi,&\text{otherwise}.
\end{cases}
\]
Thus an ordinary achievement is written without a temporal operator in the
query artifact, while both query classes enter the same LTLf2DFA/MONA path.
\end{definition}

\begin{definition}[Bound proposition valuation]
For validated bound query $\widehat\tau_q=(\tau_q,\theta_q)$, substitution
$\theta_q$ replaces declared parameters with type-compatible objects and leaves
domain constants fixed. Its proposition valuation is
\[
a\in\mathit{val}_q(s)
\Longleftrightarrow
\begin{cases}
p(\bar t)\theta_q\in s_{\mathcal P},
  & \mu_q(a)=p(\bar t),\\
s_{\mathcal F}(f(\bar t)\theta_q)=c,
  & \mu_q(a)=[f(\bar t)=c].
\end{cases}
\]
Thus predicate atoms and supported integer equalities are evaluated in the same
bound PDDL state.
\end{definition}

\begin{definition}[Finite-trace satisfaction]
Let $\xi=s_0\ldots s_n$ be a nonempty finite state trace and
$0\leq i\leq n$. Satisfaction for $\Phi_{\mathrm{syn}}$ is
\[
\begin{aligned}
\xi,i\models a
  &\Longleftrightarrow a\in\mathit{val}_q(s_i),\\
\xi,i\models\neg a
  &\Longleftrightarrow a\notin\mathit{val}_q(s_i),\\
\xi,i\models\varphi\land\psi
  &\Longleftrightarrow \xi,i\models\varphi\text{ and }\xi,i\models\psi,\\
\xi,i\models X\varphi
  &\Longleftrightarrow i<n\text{ and }\xi,i+1\models\varphi,\\
\xi,i\models F\varphi
  &\Longleftrightarrow \exists j\in[i,n]:\xi,j\models\varphi,\\
\xi,i\models \varphi U\psi
  &\Longleftrightarrow \exists j\in[i,n]:\xi,j\models\psi\text{ and }\\
  &\hspace{4.8em}\forall k\in[i,j):\xi,k\models\varphi.
\end{aligned}
\]
Write $\xi\models\varphi$ for $\xi,0\models\varphi$. The clauses make $X$
strong and require the right operand of $U$ to occur on the finite trace.
\end{definition}

\begin{definition}[Evaluation and certified-compilation scopes]
$\Phi_{\mathrm{bench}}\subset
(\Phi_{\mathrm{syn}}\setminus\Phi_{\mathrm{ach}})$ is the TEG family obtained by
instantiating the five registered templates in
Table~\ref{tab:teg-construction-profiles}. For fixed $D$ and $\mathcal M_D$,
$\Phi_{\mathrm{cert}}(D,\mathcal M_D)\subseteq\Phi_{\mathrm{syn}}$ contains
validated bound formulas whose required distance-reducing DFA obligations admit
the binding, completion-effect, preservation, progress, and resource
checks used by the compiler. Formula parsing alone therefore establishes
membership only in $\Phi_{\mathrm{syn}}$.

Before transition-certificate construction, the DFA converter expands its Boolean
transition relation into separate conjunctive transition records. The runtime
monitor retains the complete DFA transition relation.
\end{definition}

\subsection{Candidate BDI Modules}

\begin{definition}[Candidate branch]
A lifted candidate branch is
$b=\langle g_b,C_b,\beta_b,\Sigma_b,\pi_b\rangle$, where $g_b$ is one
achievement literal, $C_b$ is a conjunction of typed positive and negative
state literals, integer comparisons, and equality filters, $\beta_b$ is a
finite sequence of primitive PDDL actions and internal achievement calls,
$\Sigma_b$ is a conditional module-completion summary, and $\pi_b$ records
whether the branch came from provider evidence or action-schema construction.
\end{definition}

\begin{definition}[Conditional module-completion summary]
For candidate branch $b$, $\Sigma_b$ is a condition-indexed conservative
summary of effects observable when $\beta_b$ returns. Its Boolean projections
are $\mustadd_b$, $\mayadd_b$, and $\maydelete_b$; it additionally records any
exact constant-integer delta and keyed resource-mode transition. A must effect
holds for every feasible selected execution under its recorded condition. A may
effect contains every feasible selected completion effect and may overapproximate
it.
\end{definition}

\begin{definition}[Internal-call closure]
For any BDI plan set $\mathcal R$, internal-call closure is the predicate
\[
\operatorname{Closed}_D(\mathcal R)
\Longleftrightarrow
\begin{aligned}[t]
&\forall b\in\mathcal R,\ \forall !q(\bar Y)\text{ occurring in }\beta_b,\\
&\exists b'\in\mathcal R\text{ such that}\\
&\qquad g_{b'}\text{ type-unifies with }q(\bar Y)\\
&\qquad\text{under the type hierarchy of }D.
\end{aligned}
\]
The condition applies to every plan in $\mathcal R$ and therefore also to calls
reached transitively from another module. Primitive PDDL actions are not
achievement calls and are outside this predicate. The predicate establishes
type-compatible resolution only: it neither quantifies over reachable states
nor guarantees that a resolving branch is applicable in every such state.
\end{definition}

The context is an applicability condition, not a sequence of subgoals. Positive
context literals and static sort facts may bind variables. Negative literals
and comparisons only filter already bound variables. A plan such as
\[
+!on(X,Y):\neg clear(Y)\leftarrow !clear(Y);!on(X,Y)
\]
is therefore admitted only when the call to $clear(Y)$ is closed by another
selected module and the recursive call carries a well-founded progress proof.

\begin{definition}[Candidate certification]
For domain $D$, $\operatorname{Cert}_D(b)$ is the branch-certification predicate.
It holds exactly when every condition required by the structure of $b$ has a
valid finite witness:
\begin{enumerate}
\item every action and subgoal variable is range restricted and type compatible;
\item symbolic progression proves each primitive precondition;
\item successful completion establishes the trigger goal $g_b$;
\item every internal call has a declared, type-compatible target signature and
records an internal-call obligation for set-level selection;
\item every recursion cycle has a non-negative ranking that strictly decreases;
\item every temporary resource occupation has a finite, target-preserving
discharge path; and
\item $\Sigma_b$ conservatively contains every feasible completion add and delete
effect, including effects propagated through internal calls.
\end{enumerate}
Obligations that do not arise in $b$, such as resource discharge for a direct
producer that acquires no resource, are vacuous rather than unproved.
\end{definition}

The implemented relational rankings are either a global relation count or the
anchor-local count $\rho_{R,x}(s)=|\{y\mid R(y,x)\in s\}|$. The latter requires
$R$ to remain acyclic in reachable states; every recursive step must decrease
the count, and selected callees may not restore an $R(\cdot,x)$ fact.

\begin{table*}[htbp]
\centering
\small
\setlength{\tabcolsep}{3.0pt}
\begin{tabular}{
>{\raggedright\arraybackslash}p{0.17\textwidth}
>{\raggedright\arraybackslash}p{0.42\textwidth}
>{\raggedright\arraybackslash}p{0.32\textwidth}}
\toprule
Branch constructor & Additional acceptance obligation & Excluded failure \\
\midrule
Validated evidence macro & every macro step maps to a PDDL action schema and
symbolic execution establishes the singleton evidence target & invalid or
misbound provider evidence \\
Direct producer & one action adds the target under the checked context and
preserves its completion summary & action chosen only by predicate name \\
Acyclic regression & each open positive requirement is discharged without a
repeated normalized producer role or conflicting delete effect & arbitrary
depth cutoffs or cyclic schema search \\
Relational recursion & a non-negative relation feature strictly decreases and
an already-satisfied, direct-producer, or validated-evidence terminal branch
remains available & recursive oscillation \\
Target-preserving resource discharge & a finite non-repeating mode path restores keyed
availability while preserving the target & unreleased reusable resource \\
Ranked precondition repair & Clingo selects an acyclic call graph with strictly
decreasing natural-number dependency ranks & mutually recursive preparation
without a progress proof \\
\bottomrule
\end{tabular}
\caption{Candidate constructors and soundness obligations for atomic BDI
modules. Every selected
branch additionally satisfies typed binding, symbolic PDDL executability, and
target achievement on return; the resulting core satisfies
$\operatorname{Closed}_D(\mathcal M_D)$.}
\label{tab:compiler-obligations}
\end{table*}

\paragraph{Finite candidate-construction grammar.}
Let
\[
\begin{aligned}
\mathfrak G=\{&\text{producer},\text{regression},\text{recursion},
\text{precondition repair},\\
&\text{resource discharge},\text{numeric progress}\}.
\end{aligned}
\]
Let $\mathcal C_E$ be validated evidence macros and define
\[
\mathcal C_D(E)=
\operatorname{Inst}_D(\mathfrak G,T_D(E),T_D^{\mathbb Z}(E)).
\]
The instantiated set $\mathcal C_D(E)$ contains
direct producers, finite backward STRIPS regressions over acyclic
producer--precondition dependencies, certified relational recursion, and finite
causal resource-mode discharge. The regression construction builds on
goal-regression planning \cite{Fikes1972GeneralizedRobotPlans,Chen2026Moose}.
A regression state is an alpha-normalized tuple
of open requirements, the selected producer role, and resource modes. A path
stops before repeating such a state. Equivalent states retain only their
shortest body for the same completion summary. This establishes finiteness
without an arbitrary action-depth constant. Provider macros are validated but
never truncated by this search rule.

\begin{definition}[Evidence coverage]
A certified candidate $b$ covers evidence obligation $e$ when their triggers are
alpha equivalent and one of the following holds: their bodies and numeric
summaries are alpha equivalent; $b$ has the same body under a logically weaker
conjunctive context; or $b$ is an action-schema producer refinement with the same
$\mustadd$, no additional $\maydelete$, and equivalent numeric and keyed-resource
summaries. Prefix similarity alone is not coverage.
We denote this relation by $\operatorname{covers}_D(b,e)$.
\end{definition}

\begin{definition}[Schema-achievement coverage]
Let $\mathcal C_D^{\mathrm{nr}}(E)$ contain the nonrecursive
action-schema-derived branches. A certified branch $b$ realizes
$c\in\mathcal C_D^{\mathrm{nr}}(E)$, written $\operatorname{realizes}_D(b,c)$, when
either $b$ and $c$ have alpha-equivalent trigger, context, and body; $b$ is an
action-schema producer refinement with a weaker context and no weaker
completion contract; or $b$ is a target-preserving resource-discharge
alternative with the same achievement contract. This relation never licenses
body-prefix similarity alone.
\end{definition}

\begin{definition}[Feasible selection family]
The generated candidate set and its certified subset are
\[
\mathcal C_{D,E}=\mathcal C_E\cup\mathcal C_D(E),
\qquad
\mathcal C^{\checkmark}_{D,E}=
\{b\in\mathcal C_{D,E}\mid\operatorname{Cert}_D(b)\}.
\]
Let $\Omega_{D,E}$ contain evidence coverage, nonrecursive schema-achievement
coverage, preparation acyclicity, and recursive-ranking compatibility.
Per-branch binding, execution, achievement, completion-effect,
and resource-discharge obligations are already contained in
$\operatorname{Cert}_D$. The feasible selection family is
\[
\mathfrak F_{D,E}=
\{\mathcal M\subseteq\mathcal C^{\checkmark}_{D,E}\mid
\mathcal M\models\Omega_{D,E}\land
\operatorname{Closed}_D(\mathcal M)\}.
\]
Thus every $\mathcal M\in\mathfrak F_{D,E}$ covers every evidence rule and required
nonrecursive action-schema achievement contract, is internally closed, and
admits the selected recursion and cross-module dependency ranks.
Equivalently, schema-achievement coverage requires
\[
\forall c\in\mathcal C_D^{\mathrm{nr}}(E),\quad
\exists b\in\mathcal M:\operatorname{realizes}_D(b,c).
\]
\end{definition}

The certified atomic module core is selected directly as
\[
\mathcal M_D=
\operatorname*{lexargmin}_{\mathcal M\in\mathfrak F_{D,E}}
\operatorname{Cost}(\mathcal M),
\]
where $\operatorname{Cost}$ is the lexicographic objective in the main paper.
The core is represented by BDI trigger--context--body rules. It is
query-independent, closed under its internal achievement calls, and contains
no query identifier, DFA state, or monitor symbol. The evaluated AgentSpeak(L)
realization preserves these rules but does not define the core.

\begin{definition}[Maintained domain library]
For bound query $q$, let $\mathcal Q_q$ contain its top-level goal, DFA
dispatcher, and certified transition-repair plans. For an ordered sequence of
appended queries $q_1,\ldots,q_k$, the maintained domain library
$\mathcal L_D^{[k]}$ is
\[
\mathcal L_D^{[0]}=\mathcal M_D,
\qquad
\mathcal L_D^{[k]}=
\mathcal M_D\cup\bigcup_{i=1}^{k}\mathcal Q_{q_i}.
\]
$\mathcal M_D$ and $\mathcal Q_{q_i}$ are logical subsets of this one library, not
separately maintained plan-library artifacts. The compiler establishes
$\operatorname{Closed}_D(\mathcal L_D^{[0]})$ and accepts an append only when
$\operatorname{Closed}_D(\mathcal L_D^{[k+1]})$ holds. Consequently,
$\operatorname{Closed}_D(\mathcal L_D^{[k]})$ is an invariant of every
maintained library.
\end{definition}

GP2PL asks a Clingo answer-set optimizer \cite{Gebser2018Clingo} to compute
$\mathcal M_D$ by minimizing the declared lexicographic cost over
$\mathfrak F_{D,E}$. The result is globally optimal only over
$\mathcal C^{\checkmark}_{D,E}$; candidate generation is not claimed complete
for arbitrary planning programs.

\subsection{Explicit Assumptions and Rejection Boundary}

\begin{assumption}[Planning fragment]
Boolean actions use deterministic typed STRIPS semantics. Numeric effects are
constant integer increases or decreases over finite values. Actions are
instantaneous and states are fully observable. Arbitrary expressions,
continuous change, conditional effects, durative actions, probabilistic actions,
and partial observability are rejected.
\end{assumption}

\begin{assumption}[Evidence boundary]
Evidence is finite, singleton-goal, and schema mappable. It may be incomplete,
but a provider macro accepted as evidence must pass symbolic PDDL execution.
Negative achievement evidence without a schema-validated achieving action
sequence is rejected.
\end{assumption}

\begin{assumption}[Recursive progress]
A same-predicate recursive branch is accepted only when an action-schema-derived
ranking function $\rho_b:\operatorname{States}(D)\rightarrow\mathbb N$,
obtained from a non-negative relational feature, strictly decreases until a producer or
already-satisfied exit applies. An acyclic support-depth orientation additionally
assumes that the grounded support relation remains acyclic in reachable states.
Cross-predicate precondition repair uses a selected acyclic call graph for a
candidate core $\mathcal M$ and a dependency rank $\kappa_{\mathcal M}$ over
its trigger signatures such that
$\kappa_{\mathcal M}(g)>\kappa_{\mathcal M}(h)$ for every selected call edge
$g\rightarrow h$.
\end{assumption}

\begin{assumption}[Target-preserving resource discharge]
A reusable-resource sequence is accepted only when effect unification identifies
a stable capacity key. Its abstract resource-mode graph is
$\mathcal G_b^{\mathrm{res}}=(V_b^{\mathrm{res}},E_b^{\mathrm{res}})$, where a
vertex is an alpha-normalized keyed occupancy mode and a labelled edge
$(u,a,v)\in E_b^{\mathrm{res}}$ denotes a symbolically executable action-schema
transition from mode $u$ to mode $v$ that preserves the atomic target. The graph
must be finite, and symbolic
execution proves that the discharge removes occupation, restores availability,
and preserves the atomic target. Ambiguous capacity roles are rejected.
\end{assumption}

\begin{assumption}[Query fragment]
The input formula belongs to $\Phi_{\mathrm{syn}}$, and
$\operatorname{emb}(\varphi_q)$ is translated by the LTLf2DFA/MONA construction
\cite{Fuggitti2020LTLf2DFA,Henriksen1995MONA}. The evaluation distribution is
$\Phi_{\mathrm{bench}}$; sound query compilation additionally requires
membership in $\Phi_{\mathrm{cert}}(D,\mathcal M_D)$. Each compiled progress
guard is a conjunction of signed predicate literals and supported integer
equalities. Unsupported arithmetic is rejected.
\end{assumption}

\begin{assumption}[Primitive-step observation]
The deterministic finite automaton consumes the initial valuation and every
state produced by a successful primitive PDDL action. Only the environment can
update query-local monitor beliefs. Atomic module return is not the semantic
observation boundary of the full method.
\end{assumption}

\begin{table}[htbp]
\centering
\small
\setlength{\tabcolsep}{2pt}
\begin{tabular}{@{}
>{\raggedright\arraybackslash}p{0.22\columnwidth}
>{\raggedright\arraybackslash}p{0.31\columnwidth}
>{\raggedright\arraybackslash}p{0.39\columnwidth}@{}}
\toprule
Construct & Accepted strategy & Rejection boundary \\
\midrule
Positive predicate & Certified atomic module & No executable achieving branch \\
Numeric equality & Bounded constant-integer progress & Unsupported arithmetic or
incomplete numeric effect \\
Positive conjunction & Threat-safe serialization & Incomplete effects or cyclic
threats \\
Negative predicate & Observe absence or use an exact certified deleter & Synthetic
negated achievement or possible re-addition \\
Mixed Boolean and numeric guard & Complete net-effect and preservation certificate &
Missing net-effect or preservation witness \\
Primitive-state safety & DFA monitor after every primitive action & Completion-only
observation \\
\bottomrule
\end{tabular}
\caption{Supported compilation fragment and fail-closed boundary. Rejection
means that GP2PL emits a structured diagnostic rather than an uncertified
controller or synthetic atomic goal.}
\label{tab:supported-fragment}
\end{table}

\section{Complete Proofs}
\label{app:proofs}

\subsection{Atomic Compilation}

\begin{lemma}[Range-restricted grounding]
If the binding obligation for branch $b$ holds and its context is true under a
type-consistent grounding $\theta$, then every variable in every primitive
action and internal call of $\beta$ is grounded by $\theta$ or by a preceding
positive context witness.
\end{lemma}

\begin{proof}
The binding checker constructs the bound-variable set from trigger variables,
positive context literals, and validated static sort witnesses. A body term is
accepted only if it is a constant or belongs to this set. Negative contexts and
comparisons are checked after their arguments enter the set and never add a
binding. Induction over body positions therefore yields a ground, type-compatible
term tuple at every primitive action or call.
\end{proof}

\begin{lemma}[Symbolic progression soundness]
Let $\beta=a_1;\ldots;a_m$ be an action-only body certified executable from
context $C$. For every ground state $s\models C$ and compatible grounding
$\theta$, each $a_i\theta$ is applicable in the state reached after the prefix,
and the symbolic final facts and numeric values hold in the concrete successor.
\end{lemma}

\begin{proof}
The symbolic state initially contains every positive fact and numeric equality
required by $C\theta$. At step $i$, certification requires every positive,
negative, equality, and numeric precondition of $a_i\theta$ in that symbolic
state. The update removes delete effects, applies add effects, and applies each
constant numeric delta, exactly matching $\gamma$. Induction on $i$ proves that
the concrete and symbolic states agree on every tracked requirement. The final
symbolic requirements therefore hold concretely.
\end{proof}

\begin{lemma}[Well-founded recursive calls]
A selected recursive component satisfying either a strictly decreasing
non-negative relational ranking or strictly decreasing natural-number module
ranks cannot execute infinitely while remaining inside that component.
\end{lemma}

\begin{proof}
For same-predicate recursion, every recursive return-to-trigger step decreases
$\rho_b$ and no selected body can restore the protected portion of
the ranked relation. An infinite descent in $\mathbb N$ is impossible. For
cross-predicate precondition repair, every call edge decreases
$\kappa_{\mathcal M}$, so a
call cycle would imply
$\kappa_{\mathcal M}(g_0)>\kappa_{\mathcal M}(g_1)>\cdots>
\kappa_{\mathcal M}(g_0)$, a contradiction. Each
terminal rank has a direct producer, provider macro, or already-satisfied exit
because the corresponding nonrecursive schema-achievement contract is covered.
\end{proof}

\begin{lemma}[Finite target-preserving resource discharge]
If a resource-discharge certificate exists for a producer body, its discharge
sequence terminates, restores the keyed availability mode, removes the acquired
occupation mode, and leaves the atomic target true.
\end{lemma}

\begin{proof}
The abstract search state contains the capacity key and normalized occupation
mode. The accepted path does not repeat a mode, so finiteness of the mode graph
implies termination. Each edge is a PDDL action validated by symbolic
progression. Acceptance explicitly requires deletion of the current occupation,
restoration of the matched availability fluent, and absence of a delete effect
on the established target. Lemma 2 transfers these properties to concrete
execution.
\end{proof}

\begin{lemma}[Conservative completion summary]
For a selected closed module call, every Boolean completion effect possible
under a feasible selected branch is contained in its conditional $\mayadd$ or
$\maydelete$ summary, and every $\mustadd$ effect holds for all feasible selected
branches under the recorded condition.
\end{lemma}

\begin{proof}
Primitive branch summaries are computed by exact sequential add/delete
progression, so internally deleted-and-restored facts are not reported as final
deletes. For each recorded branch condition, the selected program contains a
finite universe of alpha-normalized effect templates. A worklist initializes
each summary with its primitive net effects, then substitutes the current
summary of every selected internal call.

Each update can only add a feasible template to a may set or remove a template
from a must set when a feasible alternative does not establish it. These two
directions are monotone and the template universe is finite, so the worklist
terminates. Ranked recursion is propagated at the same trigger anchors;
cross-module calls follow their finite dependency ranks.

Induction on a finite terminating call unfolding shows that every completion
effect of depth $k$ appears after at most $k$ propagation rounds. Lemma 3 makes
every selected unfolding finite. At convergence, union over feasible
alternatives therefore contains every may effect, while intersection over
compatible alternatives retains only must effects common to all of them.
\end{proof}

\begin{proposition}[Atomic branch soundness]
If an emitted branch for $p(\bar X)$ has a true context under a type-consistent
grounding, and each called selected module satisfies the achievement contract
and completion summary established by its certification, then the
branch is executable and successful return establishes $p(\bar X)$.
\end{proposition}

\begin{proof}
Lemma 1 grounds every body element. Consider body positions in order. Primitive
actions are executable and preserve the checked symbolic state by Lemma 2.
For an internal call, the induction hypothesis on the internally closed selected
module
establishes its goal and its conservative completion summary updates the caller's
symbolic state; Lemma 5 ensures that this update omits no feasible completion
effect. Lemma 3 excludes infinite selected recursion, and Lemma 4
completes every certified target-preserving keyed resource-mode discharge. The
branch achievement obligation requires
the grounded trigger in the final symbolic state; Lemma 2 and the call induction
therefore establish it in the concrete return state.
\end{proof}

\begin{lemma}[Answer-set encoding correspondence]
For every $\mathcal M\subseteq\mathcal C^{\checkmark}_{D,E}$, the grounded
selection program has an answer set whose selected candidates are exactly
$\mathcal M$ if and only if $\mathcal M\in\mathfrak F_{D,E}$. For every such
answer set, its prioritized optimization vector equals
$\operatorname{Cost}(\mathcal M)$.
\end{lemma}

\begin{proof}
The program contains one choice atom for each certified candidate. Its coverage
rules derive a witness exactly for the relations $\operatorname{covers}_D$ and
$\operatorname{realizes}_D$ encoded in the finite instance. Integrity
constraints reject a selection precisely when an evidence obligation, a
nonrecursive schema-achievement obligation, or an internal call lacks a
witness.

Additional rank atoms witness the selected acyclic preparation graph and the
declared recursive compatibility conditions. Thus projecting any answer set to
its choice atoms yields a member of $\mathfrak F_{D,E}$. Conversely, a feasible
set supplies every required coverage and rank witness, so extending its choice
atoms with those derived witnesses satisfies every rule and constraint.

Finally, each weak-constraint term is the corresponding component of the
declared lexicographic cost, with the same priority order. Projection therefore
preserves both feasibility and objective value.
\end{proof}

\begin{proposition}[Certified-candidate-set optimality]
If the answer-set optimizer returns $\mathcal M_D$, then $\mathcal M_D$ satisfies every
encoded evidence-coverage, schema-achievement, internal-call,
preparation-acyclicity, and recursive-ranking obligation and is
lexicographically optimal among feasible subsets of
$\mathcal C^{\checkmark}_{D,E}$.
\end{proposition}

\begin{proof}
By Lemma 6, answer sets correspond exactly to
$\mathfrak F_{D,E}$ and carry the same cost vector. Clingo's prioritized
optimization returns an answer set minimal in lexicographic priority order.
Its projected selection is therefore feasible and no other member of
$\mathfrak F_{D,E}$ has lower cost. The correspondence concerns only the
generated certified candidates and says nothing about ungenerated programs.
\end{proof}

\subsection{Query Composition}

An unbound query specification is
$\tau_q=\langle\iota_q,\varphi_q,\mu_q,\Theta_q,\Gamma_q\rangle$: identifier,
\ltlf{} formula, proposition-to-PDDL atom map, parameter type signature, and
binding constraints. Formally,
$\Theta_q:\bar X_q\rightarrow\mathcal T$ is the type signature, and a runtime
binding
$\theta_q:\bar X_q\rightarrow O_I$ satisfies $\Theta_q$ and $\Gamma_q$, and
$\widehat\tau_q=(\tau_q,\theta_q)$ is the validated bound query. Let
\[
\mathcal D_q=
\langle Q_q^{\mathrm{dfa}},2^{AP_q},\delta_q,q_q^0,F_q\rangle
\]
be the deterministic finite automaton of $\operatorname{emb}(\varphi_q)$ obtained
from the MONA-backed translation
\cite{DeGiacomoVardi2013,Fuggitti2020LTLf2DFA}. For transition guard $\chi$, let
\[
\chi=\bigwedge_{G\in P_\chi^+}G\land
\bigwedge_{N\in P_\chi^-}\neg N,
\qquad
\mathcal O_\chi=\langle P_\chi^+,P_\chi^-\rangle.
\]
$\mathcal O_\chi$ is its signed transition obligation. A preservation portfolio
$\Pi_{\chi,i}\subseteq\mathcal M_D$ assigns one or more certified branches to
positive occurrence $G_i\in P_\chi^+$. A negative literal is either observed
absent or has an exact certified deleter; it never invokes a synthetic negative
goal. The conditional definite/potential module-completion summaries follow the goal-
interference distinction used in BDI plan analysis
\cite{Thangarajah2003GoalInterference}; GP2PL derives them by symbolic PDDL
execution and propagates them through selected internal calls.

In $\mathcal D_q$, $Q_q^{\mathrm{dfa}}$ is the finite state set, $AP_q$ is the
proposition alphabet, $\delta_q$ is the deterministic transition function,
$q_q^0$ is the initial state, and $F_q$ is the accepting set. The map
$\mathit{val}_q:\operatorname{States}(D)\rightarrow 2^{AP_q}$ gives the
valuation induced by a PDDL state. Define the acceptance distance
$d_q(z)$ as the shortest directed-path length from
$z\in Q_q^{\mathrm{dfa}}$ to $F_q$, and $d_q(z)=\infty$ when no such path
exists. A certified progress edge $(q_s,\chi,q_t)$ must satisfy
$d_q(q_t)<d_q(q_s)<\infty$.

\begin{definition}[Threat-induced precedence]
For positive occurrences $G_i,G_j$, define
$G_j\prec_\chi G_i$ exactly when a feasible conditional $\maydelete$ effect of
a branch
in $\Pi_{\chi,j}$ unifies with $G_i$. The threat-induced precedence graph is
$\mathcal G_\chi=(P_\chi^+,\prec_\chi)$. The ordering constraint requires $G_j$
before $G_i$. A positive repair is inadmissible when a feasible $\mayadd$ effect
unifies with a protected $N\in P_\chi^-$.
\end{definition}

The transition certificate for $\chi$ records the selected occurrence-specific
portfolios, confirmation that their completion summaries are complete, the
induced threat relation, and the resulting serialization
$\boldsymbol\ell_\chi$. It is valid only when every positive repair preserves
earlier positive and negative obligations; failure to construct it rejects the
progress edge.

The certified serialization $\boldsymbol\ell_\chi$ is rendered as a balanced
binary transition-repair tree. A positive leaf observes or achieves its atom; a
negative leaf observes absence or invokes its exact certified deleter. The tree
indexes these occurrences, not DFA states or primitive actions. A direct linear
rendering would execute the same certified order but place all $n$ literal calls
and the completion test in one monolithic plan body, so the representation
parsed and stored for that plan, together with the intention continuation
constructed and resumed by the interpreter, would grow with the guard. Midpoint
decomposition preserves the identical left-to-right order while limiting every
generated controller plan to at most two body steps. Algorithm~\ref{alg:temporal}
constructs one such transition plan set.

\begin{algorithm}[htbp]
\caption{Compile a Certified DFA Progress Controller}
\label{alg:temporal}
\begin{algorithmic}[1]
\REQUIRE DFA source $q_s$, guard $\chi$, signed obligation $\mathcal O_\chi$,
selected atomic core $\mathcal M_D$
\ENSURE Transition-repair plan set $\mathcal R_{q_s,\chi}$ or rejection
\STATE $\boldsymbol\Sigma\leftarrow
\{\Sigma_b\mid b\in\mathcal M_D\}$
\STATE $(\boldsymbol\ell_\chi,\boldsymbol\Pi_\chi)\leftarrow
\textsc{CertifySerialization}(\mathcal O_\chi,\boldsymbol\Sigma)$
\IF{$(\boldsymbol\ell_\chi,\boldsymbol\Pi_\chi)=\bot$}
  \STATE \textbf{reject}
\ENDIF
\STATE emit entry $r_{q_s,\chi}\leftarrow !R_\chi[1,n];!c_{q_s,\chi}$
\STATE \textbf{procedure} $\textsc{Build}(i,j)$
\IF{$i=j$}
  \STATE emit the satisfied plan for signed literal $\ell_i$
  \IF{$\Pi_{\chi,i}\ne\varnothing$}
    \STATE emit the unmet plan restricted to $\Pi_{\chi,i}$
  \ENDIF
\ELSE
  \STATE $m\leftarrow\lfloor(i+j)/2\rfloor$
  \STATE emit $R_\chi[i,j]\leftarrow !R_\chi[i,m];!R_\chi[m+1,j]$
  \STATE $\textsc{Build}(i,m);\ \textsc{Build}(m+1,j)$
\ENDIF
\STATE \textbf{end procedure}; $\textsc{Build}(1,n)$
\STATE emit completion test $c_{q_s,\chi}$: return if monitor $\ne q_s$;
otherwise call $!r_{q_s,\chi}$
\RETURN $\mathcal R_{q_s,\chi}$
\end{algorithmic}
\end{algorithm}

Suppose the monitor leaves $q_s$ while a branch in $\Pi_{\chi,i}$ executes. After
that call returns, every suffix leaf either observes its literal without acting
or selects only a branch from its already certified $\Pi_{\chi,j}$. Applicability
is rechecked at invocation; an applicable call is PDDL-executable and preserves
the established signed prefix at module return, while a missing call fails
without inserting an uncertified action. Every primitive suffix action is
observed by the same DFA, so it may traverse further actual edges, including a
rejecting edge, but cannot manufacture a transition or report false acceptance.
Only after the suffix finishes does the completion test return to top-level
dispatch at the current state or replay when that state is still $q_s$. For
$n=1$, $R_\chi[1,1]$ is one leaf followed by the same completion test; balancing
does not choose the certified order or reduce primitive action count, but bounds
trigger fan-out and individual plan-body length by two, with logarithmic nesting
depth and linear work per pass.

\begin{definition}[Complete transition-repair pass]
For a certified serialization
$\boldsymbol\ell_\chi=(\ell_1,\ldots,\ell_n)$, let
\[
\begin{aligned}
R_\chi[i,i]&=\operatorname{Leaf}(\ell_i,\Pi_{\chi,i}),\\
R_\chi[i,j]&=R_\chi[i,m];R_\chi[m+1,j] && (i<j),\\
m&=\lfloor(i+j)/2\rfloor.
\end{aligned}
\]
Let $c_{q_s,\chi}$ be the source-state completion test. The complete pass is
\[
\operatorname{Pass}_{q_s,\chi}:=R_\chi[1,n]\cdot c_{q_s,\chi},
\]
where $\cdot$ denotes sequential controller composition. Each leaf
observes its signed literal or invokes its optional certified repair call. The
runtime monitor advances at every primitive action, but the completion test
checks whether the monitor still equals the source state only after
$R_\chi[1,n]$ returns.
\end{definition}

\begin{lemma}[Signed serialization invariant]
Suppose $\boldsymbol\ell_\chi$ is a topological order of the threat-induced
precedence relation, or an order equipped with per-occurrence preserving
portfolios. After
the repair of prefix $\ell_1,\ldots,\ell_i$ returns, every positive literal in the
prefix holds and every protected negative literal is absent.
\end{lemma}

\begin{proof}
The base case follows from observing or establishing $\ell_1$. Assume the invariant
for $i-1$. If $\ell_i$ is positive, its selected branch establishes it by
Proposition 1. Every feasible completion delete that could affect an earlier
positive would induce a threat edge and hence contradict the order, unless the
per-occurrence portfolio explicitly excludes that branch. Every feasible add of
a protected negative is likewise excluded. If $\ell_i$ is negative, the leaf either
observes absence or uses an exact finite deleter whose complete net effects
preserve the prefix. Thus the invariant holds by induction.
\end{proof}

\begin{lemma}[Post-exit suffix safety]
Suppose the monitor leaves source state $q_s$ during the repair call for
$\ell_i$. After that call returns, the remaining suffix of the complete
transition-repair pass can
commit only PDDL-executable actions from certified repairs, preserves the signed
serialization prefix at module-return boundaries, and induces only transitions
of the runtime DFA. An unavailable repair, illegal primitive action, or rejecting
monitor state cannot be converted into controller success.
\end{lemma}

\begin{proof}
Proceed over suffix occurrences $\ell_{i+1},\ldots,\ell_n$. A satisfied leaf commits
no action. An unmet leaf either has no repair and fails closed, or selects an
applicable certified repair. Atomic branch soundness gives PDDL executability and
establishment of that occurrence; the signed-serialization invariant excludes
completion effects that violate the established prefix. This argument depends
on the branch context and symbolic effects, not on the monitor remaining at
$q_s$. By the primitive-step observation assumption, every committed primitive action
updates the DFA from its actual current state, and no controller plan can write
that state. The suffix can therefore traverse further real edges, but no hidden
or fabricated edge. A rejecting state lacks an accepting top-level plan, so any
such trace fails rather than being relabelled successful.
\end{proof}

\begin{proposition}[Supported transition-composition soundness]
For an accepted signed transition obligation $\mathcal O_\chi$, assume every
required repair has an applicable
certified branch, every selected call terminates, and the preservation portfolios
satisfy the signed serialization invariant. At the end of a complete
transition-repair pass, the
controller retries if the monitor reports its source state and otherwise returns
to top-level dispatch at the current state. Every monitor transition during the
pass is an actual DFA edge whose complete guard is true in an observed state;
a rejecting state cannot report success.
\end{proposition}

\begin{proof}
The signed-serialization lemma establishes every repaired prefix, independently
of whether the primitive-step monitor remains at the source. If it leaves during
a repair, the post-exit suffix lemma covers the rest of the complete
transition-repair pass. After
every successful primitive action, the monitor evaluates the original complete
outgoing cubes, not the serialized approximation. When $R_\chi[1,n]$ returns,
the completion test reads the current state: it restarts the same obligation exactly
when that state is the source and otherwise returns to top-level dispatch. A
macro and the remaining suffix may cross several edges, but each is an actual
deterministic transition and a rejecting state has no accepting shortcut. No
controller statement can manufacture a state transition.
\end{proof}

\begin{proposition}[Balanced transition-repair complexity]
For $n\geq1$ signed literals and $e$ certified repair plans, midpoint splitting
produces $2n+e+2$ query-local plans, trigger fan-out at most two, at most two
body steps per generated controller plan, nesting depth $O(\log n)$, and
$O(n)$ node visits per pass while preserving the certified literal order.
\end{proposition}

\begin{proof}
There are $n$ satisfied-leaf plans, $e$ repair alternatives, $n-1$ internal
binary nodes, one transition-entry plan, and two completion-test plans, totaling
$n+e+(n-1)+1+2=2n+e+2$. Each internal trigger has one dispatch plan and each leaf
has at most the satisfied and repair alternatives counted above, so structural
dispatch fan-out is two. Midpoint division yields height
$\lceil\log_2 n\rceil$ up to constant entry and completion frames. Induction on
a contiguous range shows that its left subtree completes before its right
subtree, so an in-order traversal equals the certified sequence. Each node is
visited once per pass, hence linear work. Entry and internal plans contain two
controller calls, a repair leaf contains at most its certified call and monitor
checkpoint, and completion plans contain at most one call; hence body length is
at most two.
\end{proof}

\begin{proposition}[Monitor trace fidelity]
For concrete PDDL execution
$\xi=s_0,a_1,s_1,\ldots,a_k,s_k$, the monitor state after $s_i$ is exactly the state
obtained by running $\mathcal D_q$ on valuations
$\mathit{val}_q(s_0),\ldots,\mathit{val}_q(s_i)$.
\end{proposition}

\begin{proof}
The environment evaluates $\mathit{val}_q(s_0)$ before dispatch, obtaining
$z_0=\delta_q(q_q^0,\mathit{val}_q(s_0))$. Assume equality
after $s_{i-1}$. The environment accepts $a_i$ only when its PDDL preconditions
hold, computes $s_i=\gamma(s_{i-1},a_i)$, then replaces the monitor state by
$z_i=\delta_q(z_{i-1},\mathit{val}_q(s_i))$. Determinism gives the unique
automaton run.
Monitor symbols are query-local and cannot be effects of a PDDL action or a BDI
plan, so no other transition is possible.
\end{proof}

\begin{proposition}[Initial-acceptance trace semantics]
If the automaton is accepting after the initial valuation, the committed action
sequence is empty and its finite state trace is $\langle s_0\rangle$.
\end{proposition}

\begin{proof}
By the observation assumption, $s_0$ is consumed before a controller requests a
primitive action. Acceptance returns immediately and commits no action. A finite
trace is a nonempty sequence of states, so the corresponding trace is
$\langle s_0\rangle$. Adding a noop would append another position and is not
semantically neutral for strong Next or Until.
\end{proof}

\subsection{Numeric and Negative Obligations}

Positive numeric equality is an achievement only when symbolic execution proves
constant integer progress toward the target. Unit changes require strict
direction; a larger constant change requires the exact predecessor. Mixed
Boolean--numeric objectives require one complete final-state proof. Negated
numeric equality is observation-only unless a complete action proves a change
away from the forbidden value. These restrictions reduce numeric reasoning to
the same completion and preservation obligations as predicates; they do not
claim arbitrary numeric planning.

For numeric preparation, the ranking witness is
$\rho_{\mathrm{num}}=\langle n_{\mathrm{miss}},d_{\mathrm{target}}\rangle$,
where the first component counts unsatisfied producer preconditions and the
second is the bounded target deficit
$d_{\mathrm{target}}(s)=|f(s)-c|$ for target equality $f=c$. For source state
$q_s$, let $\mathcal W_{q_s}$ be its non-progress self-loop guards. The signed
source invariants are
\[
\mathcal I_{q_s}=\langle I_{q_s}^+,I_{q_s}^-\rangle,
\qquad I_{q_s}^{\pm}=\bigcap_{\chi\in\mathcal W_{q_s}}P_\chi^{\pm}.
\]
A cyclic
guard may instead use $b_\chi^{\mathrm{joint}}$ only when symbolic execution
proves that one complete branch establishes every obligation in
$\mathcal O_\chi$. The accepted branch is callable from every positive
occurrence it establishes and retains every query variable appearing in
$\mathcal O_\chi$.

For strong Until, literals common to all non-progress source loops are source
invariants. Every primitive prefix before progress must preserve them. This is
why primitive-step monitoring and module-return monitoring are distinct
experimental variants: the latter cannot justify an intermediate-state safety
claim.

\section{Algorithmic Correspondence}
\label{app:code-map}

Algorithm~\ref{alg:atomic} summarizes the atomic compilation pipeline whose
stages are defined in the main paper and formalized above.

\begin{algorithm}[htbp]
\caption{Compile the Certified Lifted Atomic Module Core}
\label{alg:atomic}
\begin{algorithmic}[1]
\REQUIRE PDDL domain $D$, training split $\mathcal I_{\mathrm{train}}$,
raw provider artifact $\mathcal E_{\mathrm{raw}}$
\ENSURE Certified lifted atomic module core $\mathcal M_D$ or a rejection diagnostic
\STATE $E_0\leftarrow\textsc{NormalizeEvidence}(\mathcal E_{\mathrm{raw}},D,\mathcal I_{\mathrm{train}})$
\STATE $E\leftarrow\textsc{CanonicalLift}(E_0,D)$
\STATE $T_D(E)\leftarrow\textsc{ProducibleTargets}(E,D)$
\STATE $\mathcal C_E\leftarrow\textsc{ValidateEvidenceMacros}(E,D)$
\STATE $\mathcal C_D(E)\leftarrow
\textsc{LiftSchemas}(D,\mathfrak G,T_D(E),T_D^{\mathbb Z}(E))$
\STATE $\mathcal C^{\checkmark}_{D,E}\leftarrow
\textsc{Certify}(\mathcal C_E\cup\mathcal C_D(E),D)$
\STATE $\mathfrak F_{D,E}\leftarrow
\textsc{FeasibleSubsets}(\mathcal C^{\checkmark}_{D,E},E)$
\IF{$\mathfrak F_{D,E}=\varnothing$}
  \STATE \textbf{reject}
\ENDIF
\RETURN $\mathcal M_D\leftarrow
\textsc{ClingoLexMin}(\mathfrak F_{D,E})$
\end{algorithmic}
\end{algorithm}

Table~\ref{tab:code-map} maps each theoretical construction to the public
implementation. Module docstrings repeat these references so that code and
paper can be audited in both directions.

\begin{table*}[htbp]
\centering
\small
\setlength{\tabcolsep}{4pt}
\begin{tabular}{p{0.22\textwidth}p{0.39\textwidth}p{0.31\textwidth}}
\toprule
Formal construction & Principal public source & Verification \\
\midrule
Evidence normalization and canonical lifting &
\path{src/domain_level_planning/evidence_module.py} &
Provider-name invariance, repeated-variable sharing, domain-constant
preservation, schema, arity, numeric, and macro tests \\
Certified candidate generation and feasible-core selection &
\path{src/domain_level_planning/atomic_module_synthesis.py} &
Clingo coverage, refinement, internal-call closure, ranking, release, and
renaming tests \\
Conditional module-completion summaries and threat relation &
\path{src/domain_level_planning/certified_effects.py} &
Typed unification, net effects, negative preservation, and portfolio tests \\
DFA progress composition &
\path{src/domain_level_planning/temporal_goal_appender.py} &
Conjunction, cyclic rejection, Until, negative, mixed numeric, and alpha-renaming tests \\
Balanced binary transition-repair tree &
\path{src/domain_level_planning/transition_repair_tree.py} &
Plan count, depth, fan-out, order, negative leaf, and checkpoint tests \\
Primitive-step execution and trace commit &
\path{src/evaluation/jason_runtime/runner.py} &
Jason execution, PDDL successor, numeric update, DFA monitor, trace, and VAL tests \\
\bottomrule
\end{tabular}
\caption{Correspondence between formal GP2PL constructions and their public
realizations.}
\label{tab:code-map}
\end{table*}

\subsection{Empirical Checks of the Formal Claims}

The proofs above establish the propositions for the declared formal objects.
Table~\ref{tab:claim-checks} records separate empirical checks that exercise
their implementations. These checks are falsification attempts, not substitutes
for the proofs: a passing finite test set cannot establish a universal theorem.

\begin{table*}[htbp]
\centering
\small
\setlength{\tabcolsep}{4pt}
\begin{tabular}{p{0.14\textwidth}p{0.48\textwidth}p{0.29\textwidth}}
\toprule
Formal claim & Empirical falsification target & Released evidence \\
\midrule
Proposition 1 and Lemmas 1--5 &
Reject unbound actions, cyclic producer dependencies, non-decreasing
recursion, and ambiguous resource discharge; execute accepted branches under
PDDL semantics &
Binding, preparation, progress, and resource challenge cases; 1,228
Jason/VAL-valid traces \\
Proposition 2 and Lemma 6 &
Reject uncovered evidence or schema-achievement contracts, open internal calls,
and non-optimal encoded branch subsets &
Clingo coverage, internal-call closure, refinement, and objective tests; exact selected
libraries for all 16 domains \\
Proposition 3 and Lemmas 5--6 &
Reject cyclic completion threats and forbidden negative completion additions;
preserve accepted signed guards &
Threat and negative-guard challenge cases; conjunction, negation, Until, and
mixed-numeric executions \\
Proposition 4 &
Check exact plan count, fan-out, depth, and in-order leaf visitation over
singleton and conjunctive guards &
Balanced transition-repair complexity and generated-controller tests \\
Proposition 5 &
Compare primitive-step monitor updates with an independent replay and direct
finite-trace semantics &
1,228 dual-DFA trace checks and 14 labelled semantic-conformance traces \\
Proposition 6 &
Accept an initially satisfied predicate or integer equality without inventing
a primitive action &
Two zero-action Jason and singleton-state replay cases \\
\bottomrule
\end{tabular}
\caption{Finite empirical falsification checks associated with the formal
claims. A passing check exercises the implementation but does not replace the
corresponding proof.}
\label{tab:claim-checks}
\end{table*}

\section{Data Appendix}
\label{app:data}

\subsection{Novel Temporal-Goal Dataset}
\label{app:teg-dataset}

\paragraph{Construction unit.}
For domain $D$ and held-out problem $P_i$, one private construction record is

\[
\mathcal B_i=\langle D,P_i,q_i,T_i,\theta_i,\pi_i\rangle .
\]

Here $q_i$ is the public controlled-language query, $T_i$ the hidden lifted
\ltlf{} oracle, $\theta_i$ its problem-object binding, and $\pi_i$ a legal
primitive-action witness. $P_i$ supplies typed objects and its initial state.

The original PDDL achievement goal is retained only as provenance. It cannot
affect rollout enumeration, candidate ranking, query wording, or construction
success; mechanically using $F$ of that goal would not test temporal order.

\begin{algorithm}[htbp]
\caption{Witness-backed construction of the TEG benchmark}
\label{alg:teg-benchmark-construction}
\begin{algorithmic}[1]
\REQUIRE Domain $D$ and canonically ordered held-out problems $P_1,\ldots,P_n$
\ENSURE Public rows $q_i$ and sealed oracles $(T_i,\theta_i,\pi_i)$
\STATE Parse the typed PDDL catalogue and each problem's objects and initial state.
\FOR{each $P_i$ in canonical path order}
\STATE Enumerate bounded legal, state-changing, nonrepeating rollouts $\Pi_i$.
\STATE Extract predicate-add, predicate-delete, and integer-change events.
\STATE Instantiate the five profiles in Table~\ref{tab:teg-construction-profiles}.
\STATE Reject candidates that fail legality, novelty, or trace satisfaction.
\STATE Lift objects to typed parameters; seal each binding and witness.
\STATE Select by profile use, semantic-signature use, quality, then a deterministic structural tie-break.
\STATE Render $q_i$ deterministically; emit separate public and sealed records.
\ENDFOR
\STATE Deduplicate only identical complete model inputs with agreeing semantics.
\end{algorithmic}
\end{algorithm}

\paragraph{Legal rollouts and events.}
No classical or generalized planner is called. Applicable actions satisfy all
positive, negative, numeric, and type conditions; effects are applied
simultaneously.

Only changed successor states are retained, and no path may repeat a complete
state.

A positive event changes a predicate atom from false to true; a negative event
changes it from true to false. A changed integer fluent contributes equality to
its successor value.

Table~\ref{tab:teg-construction-profiles} gives the exact candidate grammar.

\begin{table*}[htbp]
\centering
\small
\setlength{\tabcolsep}{4pt}
\begin{tabular}{p{0.20\textwidth}p{0.27\textwidth}p{0.45\textwidth}}
\toprule
Profile & Hidden lifted oracle & Required grounded witness evidence \\
\midrule
Same-state conjunction & $F(A\land B)$ &
One legal action makes distinct positive events $A$ and $B$ true in the same
successor state. \\
Same-state with negation & $F(A\land\neg B)$ &
One legal action makes $A$ true and deletes predicate atom $B$ in the same
successor state. \\
Ordered two milestones & $F(A\land X(F(B)))$ &
Two legal state-changing actions; $A$ is new after the first and $B$ is new
after the second. \\
Ordered three milestones & $F(A\land X(F(B\land X(F(C)))))$ &
Three legal state-changing actions produce four pairwise-distinct states and a
new selected event at each step. \\
Persistence Until & $A\,U\,B$ &
Dynamic predicate $A$ holds in every state before a new $B$ event, and $B$ is
absent initially. \\
\bottomrule
\end{tabular}
\caption{Version-1 temporal profiles. $F$ is eventuality, $X$ is strong next,
and $U$ is strong Until. The witness is grounded before $A$, $B$, and $C$ are
lifted.}
\label{tab:teg-construction-profiles}
\end{table*}

\paragraph{Registered bounds and nontriviality.}
Table~\ref{tab:teg-construction-bounds} fixes the finite search. The expanded
tier is tried once and only when a problem has no primary candidate; every
other rule remains unchanged.

\begin{table}[htbp]
\centering
\small
\setlength{\tabcolsep}{3pt}
\begin{tabular}{p{0.58\columnwidth}rr}
\toprule
Bound & Primary & Expanded \\
\midrule
Trace depth & 3 & 3 \\
Applicable actions retained per state & 12 & 32 \\
Join bindings retained per schema & 64 & 2,048 \\
First-action prefixes & 4 & 4 \\
Second actions per prefix & 4 & 4 \\
Three-step prefixes & 1 & 1 \\
Third actions per prefix & 4 & 4 \\
Candidates per problem & 1,024 & 1,024 \\
Candidates per profile & 16$^{a}$ & 16$^{a}$ \\
\bottomrule
\end{tabular}
\caption{Registered construction bounds. $^{a}$The ordered-three profile is
capped at eight candidates.}
\label{tab:teg-construction-bounds}
\end{table}

Candidates are rejected when an action has no state effect, a complete state
repeats, a second action restores the initial state, an ordered milestone is
not new at its selected step, or an Until right-hand milestone holds initially.

Failure means no witness was found under these bounds, not that no TEG exists.

\paragraph{Lifting and deterministic selection.}
Objects are traversed in formula order. Domain constants remain constants;
other first occurrences become $X,Y,Z,\ldots$ with their PDDL types.

Repeat occurrences reuse the same parameter. Type-compatible distinct objects
receive explicit inequality constraints.

The ground assignment is stored only in $\theta_i$. Each candidate receives an
alpha-normalized semantic signature over its profile, types, sharing,
constraints, atoms, and formula tree; ground object and witness-action names
are excluded.

Within each domain, selection minimizes prior profile use, then prior signature
use, a registered structural quality key, and a spelling-independent deterministic tie-break.

Thus an alphabetically early action schema cannot define the benchmark by being
encountered first.

\paragraph{Concrete example.}
A legal witness whose milestones are \texttt{holding(b1)} and later
\texttt{on(b1,b4)} is lifted with
$\theta_i=\{X\mapsto b1,Y\mapsto b4\}$ to the hidden oracle
$T_i=F(\mathit{holding}(X)\land X(F(\mathit{on}(X,Y))))$.

The public query exposes only typed $X,Y$, $X\ne Y$, and their strict milestone
order.

\paragraph{Rendering, deduplication, and audit boundary.}
A deterministic visitor renders controlled English directly from $T_i$; no
language model or domain-specific gloss participates.

The public manifest keeps one row per problem, while the sealed audit keeps
$T_i$, $\theta_i$, $\pi_i$, and the replayed states.

The 1,228 public rows are grouped into 475 translation calls by hashing the
rendered domain prompt, source text, typed parameters, constraints, and
parameter semantics.

Equal English alone is insufficient, and a group is rejected if its hidden
semantic signatures disagree.

The implementation is in
\path{src/temporal_input/nl_benchmark.py} and
\path{src/temporal_input/translation_worklist.py}; construction-profile,
lifting, symbol-invariance, and deduplication tests accompany both modules.

The novel GP2PL data release contains 475 deduplicated controlled-language
specifications and 1,228 bound query cases over 16 PDDL domains. Each record
contains the source utterance, typed parameters, binding, atom table, gold
parametric LTLf formula, model-produced JSON, language-equivalence result, and
sealed construction-witness checks. Per-domain JSON files, the complete
benchmark, model predictions, independent validation, and source archives are
included in the code-and-data appendix under
\path{paper_artifacts/temporal_goal_benchmark/v1}. This dataset evaluates
translation equivalence and witness validity; the 1,228 Jason, VAL, and
trace-automaton executions are reported in the separate evaluation release.
The released predictions use GPT-5.5; the model identifier and complete
request/response archives are retained.

The dataset includes its own data dictionary, citation metadata, and integrity
checker. Publication replaces three machine-local directory strings in the
model-delivery archive with release-relative paths. The release records the
original archive digest and every normalized member; model responses and
semantic-validation records are unchanged.

Original natural-language annotations, temporal specifications, model outputs,
generated plan libraries, and result records are released under CC BY 4.0. The
GP2PL source is released under Apache-2.0. Third-party PDDL files retain their
upstream terms and are reconstructed by pinned download scripts rather than
relicensed as original GP2PL data.

\subsection{Frozen Natural-Language Translation Prompt}
\label{app:translation-prompt}

The natural-language translation protocol is GP2PL's typed input contribution,
separate from the policy-lifting compiler. Its role is to map one controlled-language
specification to typed, externally bound parametric \ltlf{} JSON. It cannot
choose PDDL actions, objects, bindings, or domain-library modules. The full
protocol comprises five declared components: a normal-form guide,
schematic few-shot examples, lifted-variable rules, an operator whitelist, and
model-correctable retry guidance. The PDDL catalogue and public sample row are
deterministic substitutions. Hidden profile labels, object bindings, gold
formulas, and construction witnesses are not model-visible.

The frozen template below instantiates the proper-TEG translation study and
therefore lists its five temporal profiles. The general query contract also
admits an ordinary achievement as an atemporal literal or conjunction; such an
input retains that form and receives the completion embedding only at DFA
construction, as defined in Sec.~1.

The following is the semantically complete paper-facing template. Line wrapping
is typographic, and angle-bracket fields denote deterministic catalogue or
sample substitutions. The release retains the rendered requests, responses,
model identifier, prompt configuration, and validation records.

\begingroup
\small
\begin{verbatim}
SYSTEM
You translate one controlled
natural-language temporal goal into lifted
Linear Temporal Logic
on finite traces (LTLf).
Planning domain: <domain>
Your only task is semantic translation.
Do not plan or execute actions. Do not
ground lifted
parameters or choose problem objects.

PUBLIC PDDL CATALOGUE
Predicates:
  <name(argument types), one per line>
Numeric functions:
  <name(argument types), one per line>
Domain constants:
  <constant : type, one per line>
Type hierarchy (child <: parent):
  <child <: parent, one per line>

ATOM TABLE CONTRACT
- The formula uses only a0, a1, a2, ... .
- Define each symbol exactly once in atoms,
  in
  first-formula-occurrence order. Define no
  unused atom.
- Predicate entry keys: symbol, kind,
  predicate, args.
- Numeric-equality entry keys: symbol,
  kind, function, args, value. Value is an
  explicitly
  requested integer.
- Use only listed PDDL symbols, constants,
  parameter names, and arities.

LIFTED PARAMETER CONTRACT
- parameter_semantics externally_bound
  means a parameterized formula schema.
  Parameters
  are neither grounded objects nor LTLf
  quantifiers; an invocation supplies a
  type-safe assignment.
- Copy declared_parameters and constraints
  without changing their meaning. Preserve
  exact,
  case-sensitive parameter names.
- Every non-constant argument is one
  declared parameter. Do not invent, omit,
  rename, merge,
  or ground parameters.
- Check PDDL argument types and subtypes.
- Constraints restrict later assignments.
  Copy
  them as metadata, not LTLf propositions.

BENCHMARK-V1 OPERATOR CONTRACT
- Allowed operators are exactly
  F, X, U, &, !.
- F is eventuality; X is immediate next
  state; U is strong until and requires its
  right side.
- Parenthesize grouping explicitly.
- Do not use |, G, R, WX, implication,
  equivalence, quantifiers, or other
  operators.

BENCHMARK-V1 TEMPORAL STRUCTURES
- Same-state conjunction: F(a0 & a1).
- Positive plus negative: F(a0 & !a1).
- Ordered two: F(a0 & X(F(a1))).
- Ordered three:
  F(a0 & X(F(a1 & X(F(a2))))).
- Persistence through the first milestone:
  a0 U a1.
- "Strictly later" requires X; do not
  weaken it
  to F(a0 & F(a1)).

SCHEMATIC EXAMPLES
- "At some state, both conditions" ->
  F(a0 & a1).
- "First, then strictly later second" ->
  F(a0 & X(F(a1))).
- "First continues until second" ->
  a0 U a1.
The atom table supplies exact
current-domain
PDDL symbols.

OUTPUT CONTRACT
Return JSON only with exactly these eight
keys:
schema_version, sample_id, temporal_logic,
ltlf_formula, atoms, declared_parameters,
constraints, status.
Use schema_version 1, temporal_logic LTLf,
and
status supported.
Predicate atom:
{"symbol":"a0", "kind":"predicate",
 "predicate":"<listed name>",
 "args":["<parameter or constant>"]}
Numeric atom:
{"symbol":"a0", "kind":"numeric_equality",
 "function":"<listed name>",
 "args":["<parameter or constant>"],
 "value":1}
Copy sample_id, declared_parameters, and
constraints from the user message. Choose
only ltlf_formula and its atoms table. Add
no prose, markdown, explanations, object
names, or keys.

RETRY CONTRACT
A later user message may supply the
previous formula and one model-correctable
schema, syntax, vocabulary, arity, type,
parameter, operator, atom-table, or
semantic error. Correct only that
translation and
return the same eight-key JSON.
Infrastructure failures are not
instructions.

USER
Translate this public benchmark row. The
source text is the sole source of temporal
structure;
omitted metadata must not be inferred.
{
  "sample_id": "<sample identifier>",
  "source_text":
    "<controlled-language goal>",
  "declared_parameters":
    <typed parameters>,
  "constraints": <binding constraints>,
  "parameter_semantics": "externally_bound"
}
Return the required eight-key JSON object
now.

RETRY USER MESSAGE, ONLY AFTER A
CORRECTABLE ERROR
The previous translation failed a model-
correctable validation check:
{
  "previous_ltlf": "<previous formula>",
  "error_type": "<validator category>",
  "error_detail": "<precise diagnostic>",
  "hint": "<allowed correction>",
  "attempt": <positive integer>
}
Preserve the source query, parameters, and
constraints. Return the same eight-key JSON
only.
\end{verbatim}
\endgroup

Prompt acceptance is not the semantic oracle. The model output first passes
the exact eight-key schema, atom-table totality and one-to-one mapping, PDDL
vocabulary, arity, type,
parameter, and operator checks. It is then converted to a deterministic
automaton by LTLf2DFA and MONA and required to be finite-trace
language-equivalent to the hidden gold formula. Retries receive only
model-correctable diagnostics;
infrastructure failures never alter the requested semantics. This separation
lets the experiment evaluate prompt-based translation without trusting a
well-formed string as a correct temporal specification.

\subsection{PDDL Sources and Splits}

\begin{table*}[htbp]
\centering
\small
\setlength{\tabcolsep}{3pt}
\begin{tabular}{p{0.20\textwidth}p{0.18\textwidth}p{0.27\textwidth}p{0.27\textwidth}}
\toprule
Source & Domains & Split rule & Redistribution boundary \\
\midrule
MOOSE companion dataset, revision
\texttt{e0097051} \cite{DillonZChenMooseDataset} &
Eight classical and four numeric domains &
Upstream \texttt{training/} and \texttt{testing/} directories &
Publicly downloadable; no upstream license file detected, so fetched in place \\
Learner-policies repository, revision
\texttt{9991926f} \cite{BonetLearnerPoliciesRepository} &
Blocksworld Clear and On &
Published no-constants training and testing directories &
Publicly downloadable; no upstream license file detected, so fetched in place \\
PDDL Instances, revision
\texttt{cf19edf7} \cite{Bacchus2001AIPS00,PotasscoPDDLInstances} &
Blocksworld Tower &
First $\lfloor n/4\rfloor$ naturally sorted instances train; remainder test &
Publicly downloadable; no upstream license file detected, so fetched in place \\
D2L, revision \texttt{0620e169} \cite{RLeapD2LRepository} &
Depots &
First $\lfloor n/2\rfloor$ naturally sorted instances train; remainder test &
GPL-3.0 upstream repository; original terms retained \\
\bottomrule
\end{tabular}
\caption{Pinned public PDDL sources and deterministic split rules.}
\label{tab:data-sources}
\end{table*}

The choice of domains tests three structural sources of obligations: classical
singleton-goal evidence, bounded integer effects, and feature-definable
serialized subgoals. These are reporting groups, not domain routing classes.
All compiler decisions are derived from schema symbols, types, preconditions,
effects, evidence obligations, and selected module-completion summaries.

\section{Computational Reproducibility}
\label{app:reproducibility}

\subsection{Hardware and Software}

Experiments were conducted on an Apple M4 system with ten CPU cores and 24 GB
unified memory, running macOS 26.4.1 on arm64. No GPU was used. The fixed
software environment is Python 3.12.7, uv 0.8.19, Clingo 5.8.0
\cite{Gebser2018Clingo}, Tarski 0.9.1 \cite{RamirezFrances2020Tarski},
LTLf2DFA 1.0.2 \cite{Fuggitti2020LTLf2DFA}, MONA 1.4-18
\cite{Henriksen1995MONA}, Jason 3.1.2 \cite{Bordini2007Jason}, OpenJDK 24,
Maven 3.9.11, Docker 28.5.1, and VAL revision \texttt{3c7a1f3}
\cite{Howey2004VAL}. The Python dependency lockfile and external setup scripts
specify the software environment used by the experiments.

The external MOOSE comparison has two predeclared provenance scopes. For the
twelve original MOOSE domains, we copy only the five-model mean planning
coverage from Table~4 of arXiv \texttt{2511.11095v1}
\cite{Chen2025MooseExtended}: 1079.6 of 1080 cases. For the four GP2PL-added
domains, we execute the exact seed-specific Raw MOOSE models locally over five
seeds and validate every nonempty returned plan with VAL. The reference record
identifies the source table, its twelve per-domain means, and the disjoint
1080-case published scope and 148-case local extension. The scopes are never
selected per domain after execution.
Published MOOSE time and memory are not combined with locally measured
penalized average runtime, where a failure is charged twice the cutoff (PAR-2).

The local coverage record is frozen at
\texttt{paper\_artifacts/gp2pl\_evaluation/v1/}
\texttt{raw\_moose\_extension\_}
\texttt{five\_seed\_summary.json}. It retains 740 portable case records, the
declared five-seed protocol, and per-domain outcomes; run bookkeeping,
machine-local paths, and commands are excluded. The seed-level
valid-plan counts are 26, 23, 23, 25, and 20, giving 117 successes, 619 planner
failures, and four timeouts over 740 evaluations. Per-domain five-seed means
are 7.0/30 for Blocks Clear, 2.0/30 for Blocks On, 6.0/77 for Blocks Tower,
and 8.4/11 for Depots. All five summaries completed with zero infrastructure
failures. Coverage is interpreted on the declared case set under the recorded
resource limits and requires a VAL-valid returned plan. Runtime is not used
because the registered coverage
protocol permits differing worker isolation and forbids cross-hardware timing
comparison.

The remaining task-level references are frozen in
\texttt{external\_reference\_results.json} in the same release. The achievement
matrix contains \ExternalAchievementValidCount{}/\ExternalAchievementCaseCount{}
validated outcomes: \ExternalLAMAValidCount{}/\ExternalLAMACaseCount{} for LAMA
on classical PDDL and \ExternalMRPHJValidCount{}/\ExternalMRPHJCaseCount{} for
ENHSP MRP+HJ on numeric PDDL. Each direct temporal matrix retains all
\ExternalDirectTotalCaseCount{} identifiers. FOND4LTLf supports
\ExternalDirectSupportedCount{} and validates \ExternalDirectValidCount{};
TIDE supports \ExternalTIDESupportedCount{} and validates
\ExternalTIDEValidCount{}. Their unsupported inputs remain explicit rather than
being counted as planning failures.

The achievement-reference taxonomy is: LAMA has 591 valid plans, 249 planner
failures, 24 timeouts, and four
\texttt{no\_plan} records; MRP+HJ has 253 valid plans, 23 planner failures, and
84 timeouts. The four LAMA \texttt{no\_plan} cases have goals true in the
initial state, but the registered nonempty-plan wrapper does not credit a
zero-action solution. The reported 591/868 row is therefore conservative by
four cases.

Within the 492 inputs accepted by the direct temporal adapter, FOND4LTLf with
LAMA has 298 valid traces, 85 planner failures, and 109 compiler failures. The
736 unsupported inputs comprise 360 numeric problems and 376 identifier-
encoding rejections. The latter are an adapter boundary, not a limitation of
the \ltlf{} semantics, so the row is interpreted only on its declared supported
subset.

TIDE with LAMA accepts all 868 Boolean inputs and yields 675 valid traces, 188
\texttt{no\_plan} outcomes, and five timeouts. The remaining 360 inputs are
explicitly unsupported numeric cases: 207 contain numeric formulae and 153 use
numeric PDDL.

The achievement and FOND4LTLf matrices used 20 case workers; TIDE used 12. The
\ExternalAchievementRepairCount{} achievement and
\ExternalDirectRepairCount{} direct-temporal infrastructure failures were
rerun with one worker under the same case inputs, tool versions, limits, and
validation contract. The local and remote systems have equivalent hardware
configurations, and queue waiting is excluded from case runtime, so these
repaired records remain eligible for PAR-2. Only cases classified as
infrastructure failures are replaceable; planner failures, timeouts, unsupported
inputs, compiler failures, and validation failures remain unchanged.

\subsection{Parameters and Development Values}

MOOSE defaults and synthesis budgets follow its published algorithm and
evaluation protocol \cite{Chen2026Moose}; declared deviations are shown rather
than silently normalized.

\begin{table*}[htbp]
\centering
\small
\setlength{\tabcolsep}{3pt}
\begin{tabular}{p{0.24\textwidth}p{0.18\textwidth}p{0.18\textwidth}p{0.31\textwidth}}
\toprule
Quantity & Development values & Final value & Selection criterion \\
\midrule
MOOSE goal size & 1 & 1 & Singleton-goal evidence definition \\
MOOSE goal permutations & 3 & 3 & MOOSE Algorithm 1 default \\
MOOSE repetition seed & 0--4 & 0,1,2,3,4 &
Five independent repetitions; no pooling or best-seed choice \\
Internal MOOSE workers & 1,6,12 & 1 & Remove concurrent access to the seeded permutation stream \\
MOOSE synthesis limit & 30 minutes, 12 hours & 12 hours &
Match the MOOSE synthesis protocol rather than its test-time cap \\
MOOSE memory limit & 16 GiB & 16 GiB & Host-safe declared deviation from MOOSE's 32 GB allowance \\
Action-schema search-depth bound & finite legacy probes, none & none &
Termination by repeated normalized-state rejection \\
Compiler selector & local legacy pruning, Clingo & Clingo 5.8.0 &
Exact encoded constraints and lexicographic objective \\
Transition-repair tree branching factor & 2 & 2 & Binary structural theorem;
not empirically tuned \\
Translation temperature & 0 & 0 & Deterministic structured translation \\
Translation token limit & 60,000 & 60,000 & Accommodate the complete JSON response schema \\
Semantic retries & 0--3 & at most 3 & Retry only structured validation failure \\
Jason/VAL timeout & 30 minutes & 30 minutes each & Common per-instance evaluation cap \\
Validation workers & 6,8,10,12,16,20 & 6 paired/five-seed temporal; 20 external &
Fixed per registered experiment; not a method parameter \\
Java thread stack & default, 64 MiB & 64 MiB &
Prevent parser stack exhaustion on large query libraries \\
\bottomrule
\end{tabular}
\caption{Complete final parameters and values encountered during development.
Method-defining constants were not selected on test performance.}
\label{tab:parameters}
\end{table*}

All five MOOSE seeds use one internal worker and are compiled independently.
Seed processes may be launched concurrently in isolated operating-system
processes, but their evidence is never concatenated. The standalone five-seed
Full GP2PL study uses eight validation workers, whereas every paired atomic and
temporal ablation uses the registered six workers.

The paired temporal ablation fixes one core per domain so controller variants
receive identical inputs. The selected Certified Balanced controller is also
evaluated with each of the five independently compiled core sets using six
workers.

\subsection{Metrics, Distribution, and Validation}

The five-seed submission record is
\texttt{paper\_artifacts/gp2pl\_evaluation/v1/}
\texttt{five\_seed\_full\_compiler\_summary.json}. It retains the protocol,
all 6,140 seed--case outcomes, per-domain counts, and cross-seed outcome
patterns, so every aggregate in the tables can be recomputed directly. The five
runs contain 6,059 original-goal VAL-valid successes among
6,140 repeated evaluations. There are no timeouts or nonzero exits. Of the
1,228 distinct cases, 1,180 succeed for every seed, 46 are seed-sensitive, and
Depots \texttt{p12} and \texttt{p20} fail for every seed.

Table~\ref{tab:five-seed-atomic-domains} preserves every raw domain--seed
success count together with its cross-seed descriptive statistic.
Table~\ref{tab:domain-results} records the common seed-0 inputs used by the
paired controller ablation.

\input{sections/result_five_seed_atomic_domain_table}

\input{sections/result_domain_table}

The selected controller validates
\TemporalCrossSeedValidCount{}/\TemporalCrossSeedEvaluationCount{} executions:
all \TemporalCrossSeedQueryCount{} queries succeed for all
\TemporalCrossSeedCount{} seeds, with no seed-sensitive case. Action count is
invariant across seeds for every query. Mean per-seed PAR-2 is
\TemporalCrossSeedMeanParTwoSeconds{} $\pm$
\TemporalCrossSeedParTwoSDSeconds{} seconds. The profile decomposition,
per-domain Raw MOOSE extension results, and complete scope-separated external
references refine this aggregate result.
The profile table reports the paired-run query partition per seed; the
cross-seed selected-controller matrix repeats that partition without failures.

\input{sections/result_profile_table}

\input{sections/result_moose_reference_table}

\input{sections/result_external_reference_table}

\paragraph{Instance-planner outcomes.}
LAMA validates \ExternalLAMAValidCount{}/\ExternalLAMACaseCount{} classical
instances, while MRP+HJ validates
\ExternalMRPHJValidCount{}/\ExternalMRPHJCaseCount{} numeric instances. Their
PAR-2 values are \ExternalLAMAParTwoSeconds{} and
\ExternalMRPHJParTwoSeconds{} seconds. These rows cover disjoint fragments and
are interpreted within scope rather than as a same-instance ranking.

FOND4LTLf with LAMA supports \ExternalDirectSupportedCount{} of
\ExternalDirectTotalCaseCount{} temporal queries and validates
\ExternalDirectValidCount{}. The remaining
\ExternalDirectUnsupportedCount{} inputs are unsupported rather than planner
failures; this total includes both numeric PDDL outside the tool's scope and
identifiers rejected by the current adapter encoding. Supported-set PAR-2 is
\ExternalDirectParTwoSeconds{} seconds. Raw MOOSE extension runtime remains
omitted under its coverage-only protocol. The artifact records preserve the
planner, compiler, timeout, and unsupported outcome taxonomy.

TIDE with LAMA supports all \ExternalTIDESupportedCount{} Boolean queries and
validates \ExternalTIDEValidCount{}; its \ExternalTIDEUnsupportedCount{}
numeric inputs remain unsupported. Supported-set PAR-2 is
\ExternalTIDEParTwoSeconds{} seconds.

Diagnostic inspection localizes the variation at the method boundary. A provider
macro is a
complete lifted action sequence in MOOSE's readable evidence; a Logistics
example transports one package by truck, airplane, then truck. Seeded
singleton-goal order changes the states from which these long macros are
regressed. When one is absent, action-schema-derived candidate generation cannot always
replace it because
\texttt{at} may require \texttt{in} while loading to establish \texttt{in}
requires \texttt{at}; this is a cyclic producer dependency with no current
decreasing-mode witness. Depots \texttt{p12} and \texttt{p20} expose a
different limit: a crate may be suspended while the controller must choose and
bring a typed support resource. The current resource-discharge check proves
finite release for a capacity resource already bound in the branch, but cannot
construct a nested support-resource selection witness. A generic extension would need an
evidence-guided decreasing mode path for Logistics and a certified nested
resource choice for Depots. Domain-specific transport or parking rules are not
admissible substitutes.

All five seeded runs use the same benchmark case set and declared compiler
configuration. Timing is excluded from cross-method speed and
significance claims because the sequential runs overlapped unrelated machine
workloads. A supplementary diagnostic artifact may report the descriptive,
within-method cumulative distribution of Jason subprocess
\texttt{run\_seconds}; it excludes queue waiting and VAL time, keeps failures
in the denominator, and is not evidence of superiority over another planner.

Atomic measures are producible-predicate trigger coverage, internal-call closure,
held-out Jason plus original-goal VAL success, selected branches, context and
body cost, library size, and compilation time. Temporal success requires all
of: controller acceptance, Jason completion, neutral-goal VAL acceptance,
independent PDDL replay, gold-DFA acceptance, and predicted-DFA acceptance.
Failed action prefixes are diagnostics and never plans.

The atomic matrix repeats each of 1,228 held-out case identifiers under five
evidence seeds. We average the five paired binary differences for each identifier
and apply a two-sided exact sign test to the nonzero case-level differences;
the 6,140 seed--case outcomes are not treated as independent units.

The temporal ablation contains one query per held-out problem under a common
atomic core. Queries cluster by domain and may share one of 475 translated
inputs, so paired contrasts are descriptive. The robustness extension repeats
only the selected controller over five independent cores and reports variation
across seeds.

The completed paired release is
\texttt{paper\_artifacts/gp2pl\_evaluation/v1/}
\texttt{paired\_ablation\_results.json}. It contains 24,560 portable atomic
outcomes (1,228 cases, four variants, and five seeds), 4,912 portable temporal
ablation outcomes, \TemporalCrossSeedEvaluationCount{} selected-controller
cross-seed outcomes, and the 13 challenge outcomes.
Machine-local paths and transient logs are excluded; the report below is
computed from the recorded outcomes.

\input{sections/result_atomic_comparison_table}

Direct Producers and Maximum Feasible have
\AtomicDirectToMaximumRightOnlyValidCount{} seed--case outcomes favoring
Maximum Feasible and none favoring Direct Producers. Aggregation within case
gives 130 nonzero case-level differences, all positive; the two-sided exact
sign test yields $p=1.47\times10^{-39}$. Maximum Feasible and Full GP2PL have
identical valid-case sets; Full removes \AtomicFullBranchReduction{} mean
branches and \AtomicFullLibraryKiBReduction{} mean KiB. Evidence Only and
Direct Producers differ on one seed--case outcome in the opposite direction,
so target completion alone is not claimed to make AgentSpeak plan selection
monotone.

\input{sections/result_temporal_comparison_table}

Certified Balanced has 115 cases favoring it over Unprotected Serialization and
none favoring the unprotected controller. The four numeric domains contribute
net gains of 25, 28, 29, and 31, and Transport contributes two. Thus 113 of the
115 gains occur in the bounded-integer subset. Certified Balanced and
Module-Return Monitor have
\TemporalBalancedToModuleReturnLeftOnlyValidCount{} discordant cases, all favoring
primitive-step observation. PAR-2 and action counts remain descriptive; no
continuous-measure superiority claim is made.

Across the \TemporalCrossSeedEvaluationCount{} selected-controller executions,
runtime has minimum 1.02, first quartile 1.90, median 2.35, third quartile 4.81,
maximum 66.88, mean 5.36, and sample standard deviation 7.64 seconds.
Primitive-action count has minimum 1, first quartile 1, median 2, third quartile
2, maximum 4, mean 1.77, and sample standard deviation 0.83. The complete
records, rather than only these summaries, are included.

Thirteen registered challenge cases all pass. Seven exercise rejection or exact
acceptance for binding, cyclic preparation, obstruction exchange, ambiguous
resource capacity, cyclic transition threats, forbidden negative completion,
and exact terminal numeric progress. Six test invariance under predicate/action
renaming, action-parameter permutation, evidence-object renaming, irrelevant
static vocabulary, negative-guard renaming, and progress-relation renaming.

The completed matrix pairs binary atomic success by seed, domain, and test, and
binary temporal ablation success by query under identical inputs; the selected
controller additionally has five-seed outcomes. The atomic sign test uses
distinct case identifiers as its inferential unit. Temporal contrasts and
cross-seed summaries are descriptive. The Balanced representation claim is
structural rather than runtime-based, and continuous values carry no superiority
test.

\subsection{Reproduction Commands and Artifact Map}

The code-and-data appendix contains every GP2PL-authored source file required to
conduct and analyze the experiments: compiler and runtime source, tests,
preprocessing and benchmark-materialization programs, experiment drivers,
validation harnesses, table generators, the dependency lockfile, novel data,
fixed generated libraries, compact execution records, and manifests. The same
source is released under Apache-2.0. Third-party planners, interpreters, and
benchmark repositories are not represented as author-created code; supplied
setup scripts retrieve them at recorded revisions and preserve their upstream
terms.

\paragraph{Generator-assisted domain onboarding.}
For a future domain, a third-party PDDL problem generator may supply
$\mathcal I_{\mathrm{train}}$ and $\mathcal I_{\mathrm{test}}$ before evidence
extraction. Reproducibility requires the generator revision and digest,
parameter ranges, random seeds, deterministic split rule, and complete output
manifest with file digests.

Every generated problem must parse against $D$, training and held-out content
hashes must be disjoint, and the held-out set must be sealed before evidence
learning. The generalized-planning backend receives only
$\mathcal I_{\mathrm{train}}$; GP2PL receives its normalized evidence. A
generator supplies coverage opportunities, not certification witnesses, so every compiler
gate remains unchanged.

\begingroup
\small
\begin{verbatim}
uv sync
bash scripts/setup_mona.sh
bash scripts/setup_moose.sh
bash scripts/setup_benchmark_sources.sh
uv run python \
  scripts/materialize_achievement_\
benchmarks.py
uv run pytest -p no:cacheprovider -q
uv run python \
  scripts/run_certificate_challenge_\
matrix.py
\end{verbatim}
\endgroup

The fixed evaluation tables are regenerated with the conference-neutral
generator:

\begingroup
\small
\begin{verbatim}
RELEASE=paper_artifacts/gp2pl_evaluation/v1
uv run python \
  scripts/generate_evaluation_tables.py \
  --execution-summary \
  "$RELEASE/\
temporal_execution_summary.json" \
  --atomic-library-root \
  "$RELEASE/atomic_libraries" \
  --output-dir \
  latex_code/aamas_method_paper/sections
\end{verbatim}
\endgroup

The independent five-seed Full GP2PL table and compact public record are
regenerated directly from the five complete Jason/VAL summaries with:

\begingroup
\small
\begin{verbatim}
uv run python \
  scripts/freeze_five_seed_full_\
compiler_results.py
\end{verbatim}
\endgroup

The paired ablation release and both comparison tables are regenerated from the
completed matrix without rerunning any planner:

\begingroup
\small
\begin{verbatim}
uv run python \
  scripts/freeze_paired_ablation_\
results.py \
  --paired-results "$PAIRED_RESULTS" \
  --challenge-summary "$CHALLENGE_SUMMARY"
uv run python \
  scripts/freeze_cross_seed_temporal_\
results.py
\end{verbatim}
\endgroup

The former three-panel empirical view remains reproducible as a supplementary
diagnostic artifact generated from the frozen paired-ablation release rather
than from manually copied table values. It is not included in the main paper;
the exact endpoint results appear in the generated tables. The release itself
is derived from the complete registered runner matrix by the preceding command:

\begingroup
\small
\begin{verbatim}
RESULT=paper_artifacts/gp2pl_evaluation/\
v1/paired_ablation_results.json
uv run python \
  scripts/generate_aaai_figures.py \
  --ablation-results "$RESULT" \
  --output-file \
  latex_code/aamas_method_paper/\
figures/fig3_evaluation.png
\end{verbatim}
\endgroup

The generator emits a 4,200 by 1,700 pixel PNG at 600 dpi. Panel (a) recomputes
the four-variant atomic coverage matrix
from all five paired seeds, displays every domain on which any variant differs,
and aggregates only the 11 domains on which all variants reach 100\%. Panel (b)
uses all 385 seed--instance records from the registered Blocksworld Tower size
series, retaining failures in each cumulative denominator. Panel (c) marks 1,
10, 100, and 1,800 seconds and retains every failed temporal query in its
denominator. Generation tests reject an incomplete pairing, an inconsistent
success oracle, changed dimensions, or missing provenance.

The public release supplies an Apache-2.0 code license, a CC BY 4.0 license for
original GP2PL data, third-party notices, citation metadata, and a versioned
release of the novel temporal-goal dataset. During double-blind review, the same
files are supplied as an anonymized code-and-data appendix and the identifying
web link is disabled in the main-paper source. All external benchmark inputs are
publicly retrievable at the cited revisions. No dataset used by the experiments
remains unavailable; the checklist's universal condition concerning unavailable
datasets is therefore satisfied, rather than being used to imply that a hidden
dataset exists.

\section{Reproducibility Checklist}
\label{app:checklist-explanation}

The main checklist answers ``yes'' when the criterion is satisfied by the main
paper together with this appendix and the submitted code-and-data archive.
Specifically, this appendix supplies all formal assumptions and proofs; the
novel dataset description, complete released records, and CC BY 4.0 license;
citations, revisions, and deterministic split rules for existing public data;
all author-created preprocessing, execution, and analysis code; complete
parameters; hardware and software; distributional statistics; and a
theory-to-code map.

The availability answers distinguish authored artifacts from dependencies.
Every GP2PL-authored source file needed by the reported experiments is included
and Apache-2.0 licensed. External tools and benchmark repositories remain
third-party works: the appendix cites them, records exact revisions, and supplies
deterministic retrieval scripts without asserting ownership or relicensing them.
Likewise, the sealed construction audit was unavailable to the translation model
during inference but is included in the released temporal-goal dataset for
independent validation. Hence there is no non-public experimental dataset.

The statistical answer is grounded in the frozen paired records. Atomic
coverage inference uses a two-sided exact sign test over held-out case
identifiers after within-case aggregation across seeds. Temporal coverage,
structural equality, and continuous summaries remain descriptive. Complete
per-instance outcomes are available for independent reanalysis.
